### 2022-07-11 script/ms_report.Rmd
Add rules for variant calling with HaplotypeCaller, and report generation

---
title: "MsGenova Project: Data analysis on variant calling from Illumina data"
author: "Gabrièle Adam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
    result_dir: result_dir
    rmd: "report.Rmd"
    sample: sample
    DP.min: DP.min
    AR.min: AR.min
    region: region
    multiqc: multiqc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
```

```{r define_params, echo=FALSE, warning=FALSE, error=FALSE}
DP.min<-params$DP.min
AR.min<-params$AR.min
workspace<-params$result_dir
region<-params$region
multiqc<-params$multiqc
print(region)
```

# Source
<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>
<a download="MultiQC report" href="`r base64enc::dataURI(file = params$multiqc, mime = 'text/rmd', encoding = 'base64')`">MultiQC report in HTML formata</a>
<a download="Report_snpeff.html" href="`r base64enc::dataURI(file = params$multiqc, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on GenomicsVCF data</a>

### 2022-07-12 script/report.Rmd
Update log, test on config file, and read reference

---
title: "MultiCrispr Project: Data analysis"
author: "Gabrièle Adam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        highlight: tango 
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
    result_dir: result_dir
    rmd: "report.Rmd"
    sample: sample
    DP.min: DP.min
    AR.min: AR.min
    guide: guide
    region: region
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(vcfR)
library(ggplot2)
```

```{r define_params, echo=FALSE, warning=FALSE, error=FALSE}
DP.min<-params$DP.min
AR.min<-params$AR.min
workspace<-params$result_dir
barcodeDesign<-params$sample
guide<-params$guide
region<-params$region
```

# Introduction

# Data / Input

```{r read_input, echo=FALSE, warning=FALSE, error=FALSE}

barcode_sheet <- read.csv(file = barcodeDesign, header = TRUE, sep = ",") %>% mutate(barcodeName = paste0(barcode_R_name, "--", barcode_F_name)) 
plant.names <- barcode_sheet$plant
controls <- dplyr::filter(barcode_sheet, condition == "WT")$plant
cases    <- dplyr::filter(barcode_sheet, condition != "WT")$plant
```
* number of plants : `r length(unique(plant.names))`
* number of wells : `r length(unique(barcode_sheet$well))`
* number of conditions : `r unique(barcode_sheet$condition)`

# Analysis

## Demultiplexing results

```{r, echo=FALSE, warning=FALSE, error=FALSE}
ccs.lima.log <- read.table(paste0(workspace,"/tmp/lima_summary_extract"), sep="\t")
```
* number of CCS multiplexed reads : `r ccs.lima.log[1,1]`
* number of CCS demultiplexed reads : `r ccs.lima.log[2,1]`

## Mapping

```{r mapping_flagstat, echo=FALSE, warning=FALSE, error=FALSE,  out.width="50%", fig.show="hold"}

  file <- paste0(workspace,"/03_mapping/flagstat/Extract_data.flagstat")
  if(file.exists(file)){

    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("well", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

  }

DT::datatable(flagstat, 
              options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),
              class = 'cell-border stripe')
#count
nbr.reads.melt <- flagstat %>% dplyr::select(well, mapped, unmapped, supplementary) %>% reshape2::melt(id = "well", value.name = "nbr.reads")
summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot")
#ratio
ratio.reads.melt <- flagstat %>% dplyr::select(well, mapped.ratio, supplementary.ratio) 
names(ratio.reads.melt) <- c("well", "mapped", "supplementary")
ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "well", value.name = "ratio")
summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot2")
```

```{r coverage_analysis, echo=FALSE, warning=FALSE, error=FALSE, out.width="50%", fig.show="hold"}
bed.tab <- read.table(text=gsub(",","\t",readLines(region)))
names(bed.tab) <- c("rname", "start", "stop", "target","amplicon")
```
### Coverage analysis ontarget regions sequences {.tabset .tabset-fade}

```{r plot_coverage, results='asis',echo=TRUE, warning=FALSE, error=FALSE}

cov.list <- lapply(barcode_sheet$well, function(x){

  file <- paste0(workspace,"/03_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(well = x)
  }
})

cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "well", "nbr_reads")
cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
cov.summary <- cov %>% group_by(amplicon,target) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
cov <- left_join(cov, cov.summary, by = c("amplicon","target")) %>% mutate(x.label=paste0(target, "\nmedian=", median, "\nmax=", max, "\nmin=", min)) 
for (i in unique(bed.tab$target)){
cat("####",i,"\n")
g<-ggplot(filter(cov,target==i)) + geom_boxplot(aes(y=numreads, x=x.label)) + xlab("") + ylab("#reads") + ggtitle("Plot coverage")+facet_wrap(. ~ target,scales = "free")+theme(axis.text.x=element_text(angle=90,hjust=1))
print(g)
cat('\n\n')
}
```
</div>

```{r plot_depth}
cov.list <- lapply(barcode_sheet$well, function(x){

  file <- paste0(workspace, "/03_mapping/depth/", x, ".depth")
  if(file.exists(file)){

    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
  }
})

depth <- purrr::reduce(cov.list, rbind)
colnames(depth) <- c("rname", "position", "depth", "plant")
depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
heading<-as.character(unique(bed.tab$target))
```

### Representation for each target region by amplicon {.tabset .tabset-fade}

```{r create_fct_for_tabulation, results='asis',echo =TRUE}

for (i in unique(bed.tab$target)){
 cat("####",i,"\n")
  p <- ggplot(data = filter(depth,target== i)) + geom_line(aes(x=position, y=depth, color=plant))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~amplicon, scales = "free") 
  print(p)
cat('\n\n')
}
```

</div>

### Proportion of {.tabset .tabset-fade}

```{r proportion_of, results='asis',echo=TRUE, warning=FALSE, error=FALSE}
for (i in unique(bed.tab$target)){
 cat("####",i,"\n")
tmp <- cov  %>% filter(target== i) %>% group_by(nbr_reads, target, amplicon) %>% count(name = "#well") 
g1<-ggplot(data = tmp) + geom_bar(aes(x=amplicon, y=`#well`, fill=nbr_reads), stat = "identity") + xlab("") +facet_grid(~target, scales = "free")
g2<-cov %>% filter(target== i) %>% group_by(well, nbr_reads,target) %>% count(name = "amplicon") %>% filter(nbr_reads == ">15X") %>% group_by(amplicon,target) %>% count(name = "#well") %>% ggplot(aes(x = amplicon, y = `#well`)) + geom_bar(stat = "identity") + xlab("nbr of amplicon covered at >15X")
print(g1)
print(g2)
cat('\n\n')
}
```
</div>

## SNV and Indels calling

```{r define_r_fct_multiallele, echo=FALSE}
vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){

  vcfR.tid.tmp <- 
    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), function(x){return(length(x))}))) %>%
    dplyr::mutate(nb.allele=nb.allele-1)

  vcfR.tid.list <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){

    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
      tidyr::separate(col = ALT,   into = paste0("ALT" ,1:x), sep = ",", remove = FALSE) %>%
      tidyr::separate(col = AC,    into = paste0("AC" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = AF,    into = paste0("AF" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = MLEAF, into = paste0("MLEAF" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = MLEAC, into = paste0("MLEAC" ,1:x),  sep = ",", remove = FALSE)

    tmp.list <- lapply(1:x, FUN = function(i){

        tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
                                         gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
                                         ID_old    = ID,
                                         gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
                                         ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
                                         AC_old    = AC,    AC    = get(paste0("AC", i)),
                                         AF_old    = AF,    AF    = get(paste0("AF", i)),
                                         MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
                                         MLEAC_old = MLEAC, MLEAC = get(paste0("MLEAC", i)),
                                         ID = paste(CHROM, POS, REF, ALT, sep="_"),
                                   ) %>%
        dplyr::select(-gt_ref_AD, 
                      -paste0("ALT", 1:x), 
                      -paste0("gt_alt_AD", 1:x),  
                      -paste0("AC", 1:x),
                      -paste0("AF", 1:x),
                      -paste0("MLEAF", 1:x),
                      -paste0("MLEAC", 1:x)) %>%
        dplyr::filter(ALT != "*")

        })

        return(tmp.list %>% purrr::reduce(rbind))
    })

    return(purrr::reduce(vcfR.tid.list, rbind))
}
```

```{r , echo=FALSE, warning=FALSE, error=FALSE}

vcf_file<-paste0(workspace,"/04_snv_indels/genotype_plant.annotate.vcf")

vcfR.inst <- vcfR::read.vcfR(vcf_file)

vcfR.tid.list <- vcfR::vcfR2tidy(vcfR.inst)

vcfR.tid <- full_join(vcfR.tid.list$fix, vcfR.tid.list$gt, by = c("ChromKey", "POS")) %>% 
            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))

variants <- vcfR_split_multiAlt(vcfR.tid) %>% left_join(bed.tab, by=c("CHROM" = "rname")) %>% filter(POS >=start, POS <= stop) %>% select(-stop, -start) %>%
  dplyr::mutate(gt_AR_corr = dplyr::if_else(gt_AR < AR.min, 0, gt_AR),
         gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
         gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)),
         var_type   = dplyr::if_else(str_length(REF) == str_length(ALT), "SNV", dplyr::if_else(str_length(REF) > str_length(ALT), "DEL", "INS")),
         var_length = str_length(ALT) - str_length(REF),
         var_class  = dplyr::if_else(var_length == 0, var_type, 
                      dplyr::if_else(abs(var_length) == 1,  paste0(var_type, "1"),
                      dplyr::if_else(abs(var_length) >  50, paste0(var_type, ">50"), paste0(var_type, "<50")))))

variants.control <- variants %>% filter(Indiv %in% dplyr::filter(barcode_sheet, condition == "WT")$plant, gt_AR_corr > 0) 
variants.freq_corr <- variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr)) %>% 
  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(ID, target,amplicon) %>% count(name="freq")

variants <- variants %>% dplyr::filter(!ID %in% variants.control$ID) %>% dplyr::full_join(variants.freq_corr, by = c("ID", "target","amplicon"))
#DT::datatable(variants)
```

### nbr of variants per target and individu {.tabset .tabset-fade}

```{r plot_nbr_variant ,results='asis',echo=TRUE}
for (i in unique(bed.tab$target)){
  cat("####",i,"\n")
g<-variants  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr), target==i)  %>% 
  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(Indiv, target,amplicon) %>% count(name="nbr_variants") %>%
  ggplot() + geom_boxplot(aes(y=nbr_variants, x=amplicon))
print(g)
g2<- variants  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr), target==i)  %>%
  select(ID, Indiv, target, var_type, amplicon) %>% unique() %>% group_by(Indiv, target, var_type,amplicon) %>% count(name="nbr_variants") %>%
  ggplot() + geom_boxplot(aes(y=nbr_variants, x=var_type)) + facet_grid(~amplicon)
print(g2)
g3<- variants  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>%
  select(ID, Indiv, target, var_class,amplicon) %>% unique() %>% group_by(Indiv, target, var_class,amplicon) %>% count(name="nbr_variants") %>%
  ggplot() + geom_boxplot(aes(y=nbr_variants, x=target)) + facet_grid(amplicon~var_class) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylim(0, 10)
print(g3)
cat('\n\n')
}

```
</div>

### count per variant type and Indel length -Allele frequency in genotypes {.tabset .tabset-fade}

```{r count_allele_by_type,results='asis',echo=TRUE}
for (i in unique(bed.tab$target)){
cat("####",i,"\n")
g1<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>%  select(ID, Indiv, target) %>% unique() %>% group_by(ID, target) %>% count(name="freq") %>% ggplot() + 
	geom_bar(aes(as.factor(freq), fill=target)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(g1)
g2<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>% select(ID, Indiv, var_type) %>% unique() %>% group_by(ID, var_type) %>% count(name="freq") %>% ggplot() + geom_bar(aes(as.factor(freq), fill=var_type)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(g2)
g3<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>% select(ID, Indiv, var_class) %>% unique() %>% group_by(ID, var_class) %>% count(name="freq") %>% ggplot() + geom_bar(aes(as.factor(freq), fill=var_class)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(g3)
g4<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>% select(ID, Indiv, var_class) %>% unique() %>% group_by(ID, var_class) %>% count(name="freq") %>% ggplot() + geom_bar(aes(as.factor(freq), fill=var_class)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
cat('\n\n')
}
```
</div>

## SV Calling

```{r study_sv_calling }
vcf_file<-paste0(workspace,"/05_sv_calling/variants.pbsv.annotate.vcf")
sv.vcf <- vcfR::read.vcfR(vcf_file)

vcfR.tid.sv.list <- vcfR::vcfR2tidy(sv.vcf)

vcfR.tid.sv <- dplyr::full_join(vcfR.tid.sv.list$fix, vcfR.tid.sv.list$gt, by = c("ChromKey", "POS")) %>% 
            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))  %>%
            left_join(bed.tab, by=c("CHROM" = "rname"))%>% filter(POS >=start, POS <= stop) %>% select(-stop, -start) %>%
            tidyr::separate(col = gt_AD, into = c("gt_ref_AD", "gt_alt_AD"), sep = ",", remove = FALSE) %>%
            dplyr::mutate(gt_AR      = as.numeric(gt_alt_AD)/gt_DP,
                          gt_AR_corr = dplyr::if_else(gt_AR < 0.2, 0, gt_AR),
                          gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
                          gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)))

DT::datatable(vcfR.tid.sv)
```

### Number of variant per target {.tabset .tabset-fade}
```{r plot_sv_calling ,results='asis',echo=TRUE} 
for (i in unique(bed.tab$target)){
cat("####",i,"\n")
g1<-vcfR.tid.sv  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>% 
  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(Indiv, target,amplicon) %>% count(name="nbr_variants") %>%
  ggplot() + geom_boxplot(aes(y=nbr_variants, x=amplicon))+facet_grid(~target, scales = "free")
print(g1)
g2<-vcfR.tid.sv  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>% 
  select(ID, target,amplicon, SVLEN) %>% unique() %>% group_by(target,amplicon, SVLEN) %>% count(name="sv_length") %>%
  ggplot() + geom_boxplot(aes(y=sv_length, x=amplicon))+facet_grid(~target, scales = "free")
#print(g2)
g3<-vcfR.tid.sv  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>%
  select(ID, target, SVLEN,amplicon) %>% unique() %>% group_by(target, SVLEN,amplicon) %>%
  ggplot() + geom_point(aes(y=SVLEN, x=amplicon))+facet_grid(~target, scales = "free")
#print(g3)
cat('\n\n')
}
```
</div>

# Source
<a download="Report_snpeff_pbsv.html" href="`r base64enc::dataURI(file =params$rmd, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on pbsv data</a>
<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>
<a download="Report_snpeff.html" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on GenomicsVCF data</a>

### 2022-08-21 script/ms_report.Rmd
Add pindel rules, tdnascan rules, and beginning report

@@
 ---
-title: "MsGenova Project: Data analysis on variant calling from Illumina data"
+title: "MSGenova (Clean version) Rmakrdown Report"
 author: "Gabrièle Adam"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: "2022-08-16"
 output:
@@
 params:
+  DP.min: DP.min
+  AR.min: AR.min
   result_dir: result_dir
@@
   sample: sample
-    DP.min: DP.min
-    AR.min: AR.min
-    region: region
+  trimmo: trimmo
   multiqc: multiqc
+  vector: vector
+  flagstat: flagstat
+  haplotype_caller: haplotype_caller_vcf
 ---
@@
 library(ggplot2)
-```
-
-```{r define_params, echo=FALSE, warning=FALSE, error=FALSE}
+library(plyr)
+library(dplyr)
+library(shiny)
+vector_calling<-params$vector
+run_chromomap<-FALSE
 DP.min<-params$DP.min
@@
 workspace<-params$result_dir
-region<-params$region
-multiqc<-params$multiqc
-print(region)
+sample_sheet<-params$sample
+trimmo<- params$trimmo
+flagstat<- params$flagstat
+haplotype_caller<- params$haplotype_caller
+setwd(workspace)
 ```
 
-# Source
-<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>
-<a download="MultiQC report" href="`r base64enc::dataURI(file = params$multiqc, mime = 'text/rmd', encoding = 'base64')`">MultiQC report in HTML formata</a>
-<a download="Report_snpeff.html" href="`r base64enc::dataURI(file = params$multiqc, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on GenomicsVCF data</a>
+```{r define_function}
+search_identical_insertion<-function(final_table,threashold){
+  list_index<-c()
+  # we are gonna filter by vector, then order table bu chr, position, and sample. Next compute the distance between 2 lines (0 then che change).Next use a threshold and filter this table
+  new_tab<-as_tibble(final_table) %>% dplyr::arrange(vector,Chr,Breakpoint,sample)
+  for (i in 1:nrow(new_tab)){
+    if(i ==1){
+      #case for line =1
+      next
+    }
+    if((new_tab[i,1] == new_tab[i-1,1]) & ((new_tab[i,2] - new_tab[i-1,2]) <= threashold )){
+      list_index <- append(list_index,c(i-1,i))
+    }
+  }
+  # now get line in final table
+  identical_insertion<-new_tab[unique(list_index),]
+  return(identical_insertion)
+}
+
+freq_length_bewteen_insertion<-function(final_table){
+  # Must be done by sample
+  length_between<-c()
+  for (v in unique(final_table$sample)){
+    for (spl in unique(final_table$sample)){
+  new_tab<-as_tibble(final_table)  %>% dplyr::filter(sample==spl & vector==v) %>% dplyr::arrange(Chr,Breakpoint)
+  if(nrow(new_tab) != 0){
+  for (i in 1:nrow(new_tab)){
+    if(i ==1){
+      #case for line =1
+      next
+    }
+   if((new_tab[i,1] == new_tab[i-1,1])){
+    length_between <-append(length_between,as.integer(new_tab[i,2] - new_tab[i-1,2]))
+   }
+  }
+  }
+    }
+  }
+  return(length_between)
+}
+
+read_file<-function(pathway){
+#sample<-list.files(getwd())
+#vector<-list.files(list.files(getwd())[1])
+sample<-list.files(pathway)
+vector<-list.files(paste0(pathway,"/",sample[1]))
+i=0
+for (spl in sample){
+  for (v in vector){
+    table<-read.table(paste0(pathway,"/",spl,"/",v,"/5.",v,"_insertion.annotated.bed"),header = FALSE,col.names = c("Chr","Breakpoint","SuppRead","TDNA_info","Orientation","Freq","Annotation"))
+    if(! nrow(table)==0){
+      #add sample and vector columns
+      table["sample"]<-spl
+      table["vector"]<- v
+      if(i==0){
+        final_table<-table
+        i=i+1
+      }else{
+        final_table<-rbind(table,final_table)
+      }
+      i=i+1
+    }
+  }
+}
+return(final_table)
+}
+```
+
+`r ifelse(vector_calling,"# Study Plasmidic insertion","")`
+
+```{r define_variable}
+#sample<-list.files(getwd())
+#vector<-list.files(list.files(getwd())[1])
+pathway=paste0(workspace,"/05_tdnascan/")
+sample<-list.files(pathway)
+vector<-list.files(paste0(pathway,"/",sample[1]))
+# read sample file
+sample_sheet<-read.table(sample_sheet,sep="\t",header=TRUE)
+```
+
+```{r read_input}
+final_tab<-read_file(pathway)
+DT::datatable(final_tab)
+final_tab2<- final_tab %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector,sample)
+DT::datatable(final_tab2)
+```
+
+## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}
+
+```{r barplot_snp, results='asis'}
+#For one vector, look if it insert at the same position for several sample with a threshold for space between position
+for (v in vector){
+  sub_table<-filter(final_tab,vector==v) %>% group_by(Chr,sample) %>% dplyr::count(name="nbr_inster_by_chr")
+  #datatable(sub_table)
+  #count_tab<- count(sub_table,Chr)
+  if(nrow(sub_table)> 0){
+    cat("###",v,"\n")
+  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity")+xlab("Numbre of insertion by chromosome")+ylab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
+  print(g)
+  }
+  cat('\n\n')
+}
+```
+
+## What are the size between the insertion
+
+```{r length_between_insertion}
+list_of_size<- freq_length_bewteen_insertion(final_tab)
+#size<-as.data.frame(list_of_size,rep("insertion",times=length(list_of_size)))
+#ggplot(size,aes(x=list_of_size))+geom_density()+xlim(min(list_of_size),max(list_of_size))
+size<-as.data.frame(count(list_of_size))
+ggplot(size,aes(x=x,y=freq))+geom_point()+ylim(0,(max(size$freq)+1))
+ggplot(size,aes(x=factor(x)))+geom_bar(stat = "count")
+```
+
+## Are there the same vector insertion for several sample - New version ?
+```{r essai_shiny}
+#list_size<-freq_length_bewteen_insertion(final_tab)
+#cts<-count(list_size)
+#print(cts)
+#final_tab<-read_file("/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/05_tdnascan/data")
+essai<-search_identical_insertion(final_tab,5000)
+DT::datatable(essai)
+```
+
+`r ifelse(vector_calling,"# how many samples have insertion of which vector and which do not have any - version barplot?","")`
+
+```{r plot_by_chr_tuto, eval=TRUE, echo = TRUE}
+#Establish list of sample without any insertion
+no_insertion<-c()
+#gff3<-"~/Documents/Msgenova/ref_arabido/Ath-Col0-tair10-WG/Arabidopsis_thaliana.TAIR10.41.gff3"
+# read gff3 and extract size of chromosome ?? - > Jump to final table
+##sequence-region   1 1 30427671
+##sequence-region   2 1 19698289
+##sequence-region   3 1 23459830
+##sequence-region   4 1 18585056
+##sequence-region   5 1 26975502
+##sequence-region   Mt 1 366924
+##sequence-region   Pt 1 154478
+table_size_chr<-data.frame(c(paste0("Chr",c(1:5,"Mt","Pt"))),c(30427671,19698289,23459830,18585056,26975502,366924,154478))
+colnames(table_size_chr)<-c("Chr","Size")
+for (spl in unique(final_tab$sample)){
+  sub_table<-dplyr::filter(final_tab,sample==spl)
+  if(nrow(sub_table)==0){
+    print("yes")
+    no_insertion<-append(no_insertion,spl)
+  }else{
+ bp<- graphics::barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl),horiz=TRUE)
+ rownames(bp)<-table_size_chr$Chr
+    with(sub_table,segments(sub_table$Breakpoint,bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),bp[sub_table$Chr,]+0.5,col=sub_table$vector)
+  #print(bp)
+  }
+}
+print("List of vector without any plasmid insertion")
+print(no_insertion)
+
+```
+
+`r ifelse(run_chromomap,"#how many samples have insertion of which vector and which do not have any - version chromoMap","")`
+
+```{r chromomap , eval =run_chromomap, echo=run_chromomap}
+library(chromoMap)
+path<-"/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/chromomap/"
+chr<-paste0(path,"chr_size")
+annotation<-paste0(path,"annotation.tsv")
+for (spl in unique(final_tab$sample)){
+  sub_table<-dplyr::filter(final_tab,sample==spl)
+  if(nrow(sub_table)==0){
+    print("yes")
+    no_insertion<-append(no_insertion,spl)
+  }else{
+ bp<- barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl))
+ rownames(bp)<-table_size_chr$Chr
+  with(sub_table,segments(bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),sub_table$Breakpoint,col=sub_table$vector)
+  }}
+chromoMap(chr,annotation)
+sub_table<-dplyr::filter(final_tab,sample==spl) %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector)
+write.table(sub_table,file=paste0(path,"annot2.tsv"),sep="\t",row.names = FALSE)
+chromoMap(chr,paste0(path,"annot2.tsv"))
+```
+
+## Trimming parameters
+
+```{r trimmomatic}
+dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
+DT::datatable(dat)
+```
+
+## Mapping parameters
+
+### Flagstat results
+
+```{r flagstat}
+library(tidyverse)
+file <- flagstat
+if(file.exists(file)){
+    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
+      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
+                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
+                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
+                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
+    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")
+
+  }
+
+DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
+#count
+nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
+summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
+nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
+ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
+#ratio
+ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
+names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
+ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
+summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
+ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
+ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()
+
+```
+
+### Coverage analysis
+```{r coverage}
+cov.list <- lapply(sample_sheet$sample, function(x){
+  file <- paste0(workspace,"/03_mapping/coverage/", x, ".coverage")
+  if(file.exists(file)){
+
+    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
+  }
+})
+
+cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
+colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
+#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
+cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
+cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min)) 
+for (i in unique(sample_sheet$sample)){
+cat("####",i,"\n")
+g<-ggplot(filter(cov,sample==i)) + geom_boxplot(aes(y=numreads, x=x.label)) + xlab("") + ylab("#reads") + ggtitle("Plot coverage")+facet_wrap(.
+~ sample,scales = "free")+theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()
+print(g)
+cat('\n\n')
+}
+```
+
+#### Depth analysis
+```{r depth_analysis}
+cov.list <- lapply(sample_sheet$sample, function(x){
+
+  file <- paste0(workspace,"/03_mapping/depth/", x, ".depth")
+  if(file.exists(file)){
+    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
+  }
+})
+
+depth <- purrr::reduce(cov.list, rbind)
+colnames(depth) <- c("rname", "position", "depth", "plant")
+#depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
+heading<-as.character(unique(sample_sheet$sample))
+for (i in unique(cov$rname)){
+ cat("####",i,"\n")
+  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=plant))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free")
+  print(p)
+cat('\n\n')
+}
+
+```
+
+```{r proportion_of, results='asis',echo=FALSE, eval=FALSE, warning=FALSE, error=FALSE,results=FALSE}
+for (i in unique(bed.tab$target)){
+ cat("####",i,"\n")
+tmp <- cov  %>% filter(target== i) %>% group_by(nbr_reads, target, amplicon) %>% count(name = "#well") 
+g1<-ggplot(data = tmp) + geom_bar(aes(x=amplicon, y=`#well`, fill=nbr_reads), stat = "identity") + xlab("") +facet_grid(~target, scales = "free")
+g2<-cov %>% filter(target== i) %>% group_by(well, nbr_reads,target) %>% count(name = "amplicon") %>% filter(nbr_reads == ">15X") %>% group_by(amplicon,target) %>% count(name = "#well") %>% ggplot(aes(x = amplicon, y = `#well`)) + geom_bar(stat = "identity") + xlab("nbr of amplicon covered at >15X")
+print(g1)
+print(g2)
+cat('\n\n')
+}
+```
+
+## Study of SNV and Indels calling
+
+```{r define_r_fct_multiallele, echo=FALSE}
+vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){
+
+  vcfR.tid.tmp <- 
+    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), function(x){return(length(x))}))) %>%
+    dplyr::mutate(nb.allele=nb.allele-1)
+
+  vcfR.tid.list <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){
+
+    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
+      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = ALT,   into = paste0("ALT" ,1:x), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = AC,    into = paste0("AC" ,1:x),  sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = AF,    into = paste0("AF" ,1:x),  sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = MLEAF, into = paste0("MLEAF" ,1:x),  sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = MLEAC, into = paste0("MLEAC" ,1:x),  sep = ",", remove = FALSE)
+
+    tmp.list <- lapply(1:x, FUN = function(i){
+
+        tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
+                                         gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
+                                         ID_old    = ID,
+                                         gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
+                                         ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
+                                         AC_old    = AC,    AC    = get(paste0("AC", i)),
+                                         AF_old    = AF,    AF    = get(paste0("AF", i)),
+                                         MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
+                                         MLEAC_old = MLEAC, MLEAC = get(paste0("MLEAC", i)),
+                                         ID = paste(CHROM, POS, REF, ALT, sep="_"),
+                                   ) %>%
+        dplyr::select(-gt_ref_AD, 
+                      -paste0("ALT", 1:x), 
+                      -paste0("gt_alt_AD", 1:x),  
+                      -paste0("AC", 1:x),
+                      -paste0("AF", 1:x),
+                      -paste0("MLEAF", 1:x),
+                      -paste0("MLEAC", 1:x)) %>%
+        dplyr::filter(ALT != "*")
+
+        })
+
+        return(tmp.list %>% purrr::reduce(rbind))
+    })
+
+    return(purrr::reduce(vcfR.tid.list, rbind))
+}
+```
+
+```{r , echo=FALSE, warning=FALSE, error=FALSE}
+vcfR.inst <- vcfR::read.vcfR(haplotype_caller)
+
+vcfR.tid.list <- vcfR::vcfR2tidy(vcfR.inst)
+
+vcfR.tid <- full_join(vcfR.tid.list$fix, vcfR.tid.list$gt, by = c("ChromKey", "POS")) %>% 
+            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))
+
+variants <- vcfR_split_multiAlt(vcfR.tid) %>%
+  dplyr::mutate(gt_AR_corr = dplyr::if_else(gt_AR < AR.min, 0, gt_AR),
+         gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
+         gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)),
+         var_type   = dplyr::if_else(str_length(REF) == str_length(ALT), "SNV", dplyr::if_else(str_length(REF) > str_length(ALT), "DEL", "INS")),
+         var_length = str_length(ALT) - str_length(REF),
+         var_class  = dplyr::if_else(var_length == 0, var_type, 
+                      dplyr::if_else(abs(var_length) == 1,  paste0(var_type, "1"),
+                      dplyr::if_else(abs(var_length) >  50, paste0(var_type, ">50"), paste0(var_type, "<50")))))
+
+variants.control <- variants %>% filter(Indiv %in% dplyr::filter(sample_sheet, condition == "WT")$plant, gt_AR_corr > 0) 
+variants.freq_corr <- variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr)) %>% 
+  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(ID, target,amplicon) %>% count(name="freq")
+
+variants <- variants %>% dplyr::filter(!ID %in% variants.control$ID) %>% dplyr::full_join(variants.freq_corr, by = c("ID", "target","amplicon"))
+#DT::datatable(variants)
+```

### 2022-08-21 script/tdnaAnnot.py

@@ -30,7 +30,8 @@ def processGFF(file):
     geneInfo = {} # chr = [start, end, geneID]
     FileIN = open(file,"rU")
     for line in FileIN:
-        if "##" in line:
+        if line.startswith("#"):
+        #if "##" in line:
             continue
         else:
             data = line.strip().split("\t")

### 2022-09-13 script/ms_report.Rmd
update advanced_config file, update Genotype_vcf rule, add comments to rules

@@
 library(dplyr)
-library(shiny)
 vector_calling<-params$vector

### 2022-09-13 script/ms_report.Rmd
correct fogotten config, add yaml library to report

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+workspace<-params$result_dir
+knitr::opts_chunk$set(echo = TRUE,root.dir=workspace)
 ```
@@
 library(dplyr)
+library(yaml)
 vector_calling<-params$vector
 run_chromomap<-FALSE
+
 DP.min<-params$DP.min
 AR.min<-params$AR.min
-workspace<-params$result_dir
-sample_sheet<-params$sample
+#workspace<-params$result_dir
+
+sheet<-params$sample
 trimmo<- params$trimmo
@@
 haplotype_caller<- params$haplotype_caller
-setwd(workspace)
+#setwd(workspace)
 ```
@@
 read_file<-function(pathway){
-#sample<-list.files(getwd())
-#vector<-list.files(list.files(getwd())[1])
 sample<-list.files(pathway)
@@
 ```{r define_variable}
-#sample<-list.files(getwd())
-#vector<-list.files(list.files(getwd())[1])
 pathway=paste0(workspace,"/05_tdnascan/")
 sample<-list.files(pathway)
+print(sample)
 vector<-list.files(paste0(pathway,"/",sample[1]))
+print(vector)
 # read sample file
-sample_sheet<-read.table(sample_sheet,sep="\t",header=TRUE)
+sample_sheet<-read.table(paste0(workspace,"/genome/sample_sheet"),sep="\t",header=TRUE)
 ```

### 2022-09-13 script/ms_report.Rmd
update report, and change writing of config file, update report rule

@@
   rmd: "report.Rmd"
-  sample: sample
-  trimmo: trimmo
-  multiqc: multiqc
-  vector: vector
-  flagstat: flagstat
-  haplotype_caller: haplotype_caller_vcf
 ---
@@
 ```{r setup, include=FALSE}
-workspace<-params$result_dir
-knitr::opts_chunk$set(echo = TRUE,root.dir=workspace)
+workspace<-"/home/gabriele.adam/Documents/Msgenova/sorties_report/results"
+knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)
 ```
@@
 vector_calling<-params$vector
-run_chromomap<-FALSE
 
-DP.min<-params$DP.min
-AR.min<-params$AR.min
-#workspace<-params$result_dir
+DP.min<-5
+AR.min<-0.2
 
@@
 haplotype_caller<- params$haplotype_caller
-#setwd(workspace)
+
+config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
+config_ad<-read_yaml(config_advanced_file)
 ```
@@
 }
-return(final_table)
+if (exists("final_table")){
+return(final_table)}
+else{return(table)}
 }
@@
 ```{r define_variable}
-pathway=paste0(workspace,"/05_tdnascan/")
+pathway=paste0(workspace,"/05_tdnascan")
 sample<-list.files(pathway)
@@
 DT::datatable(final_tab)
+if(nrow(final_tab) >0){
 final_tab2<- final_tab %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector,sample)
 DT::datatable(final_tab2)
+presence_of_insertion<-TRUE
+}else{
+  presence_of_insertion<-FALSE
+}
 ```
 
-## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}
+`r ifelse(presence_of_insertion,"## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}","")`
 
-```{r barplot_snp, results='asis'}
+```{r barplot_snp, results='asis',eval=presence_of_insertion}
 #For one vector, look if it insert at the same position for several sample with a threshold for space between position
@@
     cat("###",v,"\n")
-  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity")+xlab("Numbre of insertion by chromosome")+ylab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
+  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity",position=position_dodge())+ylab("Numbre of insertion by chromosome")+xlab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
   print(g)
@@
 
-## What are the size between the insertion
-
-```{r length_between_insertion}
-list_of_size<- freq_length_bewteen_insertion(final_tab)
-#size<-as.data.frame(list_of_size,rep("insertion",times=length(list_of_size)))
-#ggplot(size,aes(x=list_of_size))+geom_density()+xlim(min(list_of_size),max(list_of_size))
-size<-as.data.frame(count(list_of_size))
-ggplot(size,aes(x=x,y=freq))+geom_point()+ylim(0,(max(size$freq)+1))
-ggplot(size,aes(x=factor(x)))+geom_bar(stat = "count")
-```
-
-## Are there the same vector insertion for several sample - New version ?
-```{r essai_shiny}
-#list_size<-freq_length_bewteen_insertion(final_tab)
-#cts<-count(list_size)
-#print(cts)
-#final_tab<-read_file("/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/05_tdnascan/data")
+# Does several samples have the same insertions ?
+```{r same_insertions, eval=presence_of_insertion}
 essai<-search_identical_insertion(final_tab,5000)
@@
 
-`r ifelse(vector_calling,"# how many samples have insertion of which vector and which do not have any - version barplot?","")`
+`r ifelse(vector_calling,"# Barplot representation of the insertions in each sample","")`
 
-```{r plot_by_chr_tuto, eval=TRUE, echo = TRUE}
+```{r plot_by_chr_tuto, eval=presence_of_insertion, echo = presence_of_insertion}
 #Establish list of sample without any insertion
@@
 }
+```
+
+```{r no_insertions, eval=presence_of_insertion}
 print("List of vector without any plasmid insertion")
@@
 
-`r ifelse(run_chromomap,"#how many samples have insertion of which vector and which do not have any - version chromoMap","")`
-
-```{r chromomap , eval =run_chromomap, echo=run_chromomap}
-library(chromoMap)
-path<-"/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/chromomap/"
-chr<-paste0(path,"chr_size")
-annotation<-paste0(path,"annotation.tsv")
-for (spl in unique(final_tab$sample)){
-  sub_table<-dplyr::filter(final_tab,sample==spl)
-  if(nrow(sub_table)==0){
-    print("yes")
-    no_insertion<-append(no_insertion,spl)
-  }else{
- bp<- barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl))
- rownames(bp)<-table_size_chr$Chr
-  with(sub_table,segments(bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),sub_table$Breakpoint,col=sub_table$vector)
-  }}
-chromoMap(chr,annotation)
-sub_table<-dplyr::filter(final_tab,sample==spl) %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector)
-write.table(sub_table,file=paste0(path,"annot2.tsv"),sep="\t",row.names = FALSE)
-chromoMap(chr,paste0(path,"annot2.tsv"))
-```
-
 ## Trimming parameters
@@
 ```{r trimmomatic}
+trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
 dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
@@
 library(tidyverse)
-file <- flagstat
+file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
 if(file.exists(file)){
@@
 cov.list <- lapply(sample_sheet$sample, function(x){
-  file <- paste0(workspace,"/03_mapping/coverage/", x, ".coverage")
+  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
   if(file.exists(file)){
@@
 
-#### Depth analysis
+### Depth analysis
 ```{r depth_analysis}
@@
 
-  file <- paste0(workspace,"/03_mapping/depth/", x, ".depth")
+  file <- paste0(workspace,"/02_mapping/depth/", x, ".depth")
   if(file.exists(file)){
@@
 depth <- purrr::reduce(cov.list, rbind)
-colnames(depth) <- c("rname", "position", "depth", "plant")
+colnames(depth) <- c("rname", "position", "depth", "sample")
 #depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
@@
  cat("####",i,"\n")
-  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=plant))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free")
+  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=sample))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free")
   print(p)
@@
 ```{r , echo=FALSE, warning=FALSE, error=FALSE}
+haplotype_caller<-paste0(workspace,"/03_snv_indels_calling/",config_ad$project_name,"_snv_indel.vcf")
 vcfR.inst <- vcfR::read.vcfR(haplotype_caller)

### 2022-09-14 script/ms_report.Rmd
update report and add extract_flagstat rule

@@
 ```{r setup, include=FALSE}
-workspace<-"/home/gabriele.adam/Documents/Msgenova/sorties_report/results"
+workspace<-params$result_dir
 knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)

### 2022-09-30 script/report.Rmd
change report for empty one

@@
 ---
-title: "MultiCrispr Project: Data analysis"
+title: "MSGenova Rmakrdown Report"
 author: "Gabrièle Adam"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: "2022-08-16"
 output:
@@
 params:
+  DP.min: DP.min
+  AR.min: AR.min
   result_dir: result_dir
   rmd: "report.Rmd"
-    sample: sample
-    DP.min: DP.min
-    AR.min: AR.min
-    guide: guide
-    region: region
+  AR: AR
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+workspace<-params$result_dir
+knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)
 ```
@@
 library(DT)
-library(vcfR)
 library(ggplot2)
+library(plyr)
+library(dplyr)
+library(yaml)
+
+config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
+config_ad<-read_yaml(config_advanced_file)
 ```
+# Tools used
 
-```{r define_params, echo=FALSE, warning=FALSE, error=FALSE}
-DP.min<-params$DP.min
-AR.min<-params$AR.min
-workspace<-params$result_dir
-barcodeDesign<-params$sample
-guide<-params$guide
-region<-params$region
-```
-
-# Introduction
-
-# Data / Input
-
-```{r read_input, echo=FALSE, warning=FALSE, error=FALSE}
-
-barcode_sheet <- read.csv(file = barcodeDesign, header = TRUE, sep = ",") %>% mutate(barcodeName = paste0(barcode_R_name, "--", barcode_F_name)) 
-plant.names <- barcode_sheet$plant
-controls <- dplyr::filter(barcode_sheet, condition == "WT")$plant
-cases    <- dplyr::filter(barcode_sheet, condition != "WT")$plant
-```
-* number of plants : `r length(unique(plant.names))`
-* number of wells : `r length(unique(barcode_sheet$well))`
-* number of conditions : `r unique(barcode_sheet$condition)`
-
-# Analysis
-
-## Demultiplexing results
-
-```{r, echo=FALSE, warning=FALSE, error=FALSE}
-ccs.lima.log <- read.table(paste0(workspace,"/tmp/lima_summary_extract"), sep="\t")
-```
-* number of CCS multiplexed reads : `r ccs.lima.log[1,1]`
-* number of CCS demultiplexed reads : `r ccs.lima.log[2,1]`
-
-## Mapping
-
-```{r mapping_flagstat, echo=FALSE, warning=FALSE, error=FALSE,  out.width="50%", fig.show="hold"}
-
-  file <- paste0(workspace,"/03_mapping/flagstat/Extract_data.flagstat")
-  if(file.exists(file)){
-
-    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
-      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
-                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
-                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
-                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
-    names(flagstat) <- c("well", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")
-
+```{r tools_used}
+space=paste0(workspace,"/00_logs/envs")
+for (file in list.files(space)){
+  env<-read_yaml(paste0(space,"/",file))
+  for (envi in env$dependencies){
+    line=sub("=="," version ",envi)
+    cat(line,"\n")
   }
-
-DT::datatable(flagstat, 
-              options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),
-              class = 'cell-border stripe')
-#count
-nbr.reads.melt <- flagstat %>% dplyr::select(well, mapped, unmapped, supplementary) %>% reshape2::melt(id = "well", value.name = "nbr.reads")
-summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
-nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
-ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot")
-#ratio
-ratio.reads.melt <- flagstat %>% dplyr::select(well, mapped.ratio, supplementary.ratio) 
-names(ratio.reads.melt) <- c("well", "mapped", "supplementary")
-ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "well", value.name = "ratio")
-summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
-ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
-ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot2")
-```
-
-```{r coverage_analysis, echo=FALSE, warning=FALSE, error=FALSE, out.width="50%", fig.show="hold"}
-bed.tab <- read.table(text=gsub(",","\t",readLines(region)))
-names(bed.tab) <- c("rname", "start", "stop", "target","amplicon")
-```
-### Coverage analysis ontarget regions sequences {.tabset .tabset-fade}
-
-```{r plot_coverage, results='asis',echo=TRUE, warning=FALSE, error=FALSE}
-
-cov.list <- lapply(barcode_sheet$well, function(x){
-
-  file <- paste0(workspace,"/03_mapping/coverage/", x, ".coverage")
-  if(file.exists(file)){
-
-    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(well = x)
-  }
-})
-
-cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
-colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "well", "nbr_reads")
-cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
-cov.summary <- cov %>% group_by(amplicon,target) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
-cov <- left_join(cov, cov.summary, by = c("amplicon","target")) %>% mutate(x.label=paste0(target, "\nmedian=", median, "\nmax=", max, "\nmin=", min)) 
-for (i in unique(bed.tab$target)){
-cat("####",i,"\n")
-g<-ggplot(filter(cov,target==i)) + geom_boxplot(aes(y=numreads, x=x.label)) + xlab("") + ylab("#reads") + ggtitle("Plot coverage")+facet_wrap(.
-~ target,scales = "free")+theme(axis.text.x=element_text(angle=90,hjust=1))
-print(g)
-cat('\n\n')
-}
-```
-</div>
-
-```{r plot_depth}
-cov.list <- lapply(barcode_sheet$well, function(x){
-
-  file <- paste0(workspace, "/03_mapping/depth/", x, ".depth")
-  if(file.exists(file)){
-
-    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
-  }
-})
-
-depth <- purrr::reduce(cov.list, rbind)
-colnames(depth) <- c("rname", "position", "depth", "plant")
-depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
-heading<-as.character(unique(bed.tab$target))
-```
-
-### Representation for each target region by amplicon {.tabset .tabset-fade}
-
-```{r create_fct_for_tabulation, results='asis',echo =TRUE}
-
-for (i in unique(bed.tab$target)){
- cat("####",i,"\n")
-  p <- ggplot(data = filter(depth,target== i)) + geom_line(aes(x=position, y=depth, color=plant))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~amplicon, scales = "free")
-  print(p)
-cat('\n\n')
 }
@@
 
-</div>
+# Trimming
+trimmomatic PE `r config_ad$params_trimmomatic ` input_r1 input_r2 output_r1 output_r2 TRAILING:3 2> log
 
-### Proportion of {.tabset .tabset-fade}
+# Alignments
 
-```{r proportion_of, results='asis',echo=TRUE, warning=FALSE, error=FALSE}
-for (i in unique(bed.tab$target)){
- cat("####",i,"\n")
-tmp <- cov  %>% filter(target== i) %>% group_by(nbr_reads, target, amplicon) %>% count(name = "#well") 
-g1<-ggplot(data = tmp) + geom_bar(aes(x=amplicon, y=`#well`, fill=nbr_reads), stat = "identity") + xlab("") +facet_grid(~target, scales = "free")
-g2<-cov %>% filter(target== i) %>% group_by(well, nbr_reads,target) %>% count(name = "amplicon") %>% filter(nbr_reads == ">15X") %>% group_by(amplicon,target) %>% count(name = "#well") %>% ggplot(aes(x = amplicon, y = `#well`)) + geom_bar(stat = "identity") + xlab("nbr of amplicon covered at >15X")
-print(g1)
-print(g2)
-cat('\n\n')
-}
-```
-</div>
+## alignments
+bwa mem -R "@RG\tID:sample\tSM:sample\tLB:Illumina\tPL:Illumina\tPU:sample" -M `r config_ad$params_bwa` reference input_fq1 input_fq2 > output
 
-## SNV and Indels calling
+## convert sam to bam
+picard SortSam --INPUT input --OUTPUT output --SORT_ORDER coordinate `r config_ad$params_sort_sam`
 
-```{r define_r_fct_multiallele, echo=FALSE}
-vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){
+## mark duplicates
+picard MarkDuplicates --INPUT input --OUTPUT{output -M metrics `r config_ad$params_mark_duplicate`
 
-  vcfR.tid.tmp <- 
-    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), function(x){return(length(x))}))) %>%
-    dplyr::mutate(nb.allele=nb.allele-1)
+# QC bam
 
-  vcfR.tid.list <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){
+## flagstat
+samtools flagstat `r config_ad$params_flagstat ` input > output
 
-    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
-      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = ALT,   into = paste0("ALT" ,1:x), sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = AC,    into = paste0("AC" ,1:x),  sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = AF,    into = paste0("AF" ,1:x),  sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = MLEAF, into = paste0("MLEAF" ,1:x),  sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = MLEAC, into = paste0("MLEAC" ,1:x),  sep = ",", remove = FALSE)
+## samtools coverage
+samtools coverage -o output `r config_ad$params_samtools_coverage ` input
 
-    tmp.list <- lapply(1:x, FUN = function(i){
+## samtools depth
+samtools depth `r config_ad$params_samtools_depth`  input > output
 
-        tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
-                                         gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
-                                         ID_old    = ID,
-                                         gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
-                                         ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
-                                         AC_old    = AC,    AC    = get(paste0("AC", i)),
-                                         AF_old    = AF,    AF    = get(paste0("AF", i)),
-                                         MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
-                                         MLEAC_old = MLEAC, MLEAC = get(paste0("MLEAC", i)),
-                                         ID = paste(CHROM, POS, REF, ALT, sep="_"),
-                                   ) %>%
-        dplyr::select(-gt_ref_AD, 
-                      -paste0("ALT", 1:x), 
-                      -paste0("gt_alt_AD", 1:x),  
-                      -paste0("AC", 1:x),
-                      -paste0("AF", 1:x),
-                      -paste0("MLEAF", 1:x),
-                      -paste0("MLEAC", 1:x)) %>%
-        dplyr::filter(ALT != "*")
+# SV Calling
 
-        })
+## Run haplotyCallze
+gatk HaplotypeCaller -R reference -I input -O output -L region --native-pair-hmm-threads 16 `r config_ad$params_haplotype_caller`
 
-        return(tmp.list %>% purrr::reduce(rbind))
-    })
+## Run GenomicsDB
+gatk GenomicsDBImport variant --genomicsdb-workspace-path output --intervals {params.region} `r config_ad$params_genomics_db`
 
-    return(purrr::reduce(vcfR.tid.list, rbind))
-}
-```
+## Run GenotypeGVCF
+gatk GenotypeGVCFs --reference reference -V gendb://input -O output `r config_ad$params_genotype_gvcf`
 
-```{r , echo=FALSE, warning=FALSE, error=FALSE}
+## Merge gvcf
+picard GatherVcfs list_of_vcf_files O=output `r config_ad$params_gathervcf`
 
-vcf_file<-paste0(workspace,"/04_snv_indels/genotype_plant.annotate.vcf")
+## Build snpeff database
+snpEff build -c output `r config_ad$params_snpeff_build` -gff3 -v reference -dataDir .
 
-vcfR.inst <- vcfR::read.vcfR(vcf_file)
+## Annotate variant
+snpEff eff -c input `r config_ad$params_snpeff_ann` -dataDir . reference -s output.html input > output
 
-vcfR.tid.list <- vcfR::vcfR2tidy(vcfR.inst)
+# Run pindel
+## run pindel
+pindel `r config_ad$params_run_pindel`  -f reference -i input -o output
 
-vcfR.tid <- full_join(vcfR.tid.list$fix, vcfR.tid.list$gt, by = c("ChromKey", "POS")) %>% 
-            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))
+## convert pindel to vcf
+pindel2vcf `r config_ad$params_convert_pindel` -r reference -P sample name -R xxx -d xx -v output
 
-variants <- vcfR_split_multiAlt(vcfR.tid) %>% left_join(bed.tab, by=c("CHROM" = "rname")) %>% filter(POS >=start, POS <= stop) %>% select(-stop, -start) %>%
-  dplyr::mutate(gt_AR_corr = dplyr::if_else(gt_AR < AR.min, 0, gt_AR),
-         gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
-         gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)),
-         var_type   = dplyr::if_else(str_length(REF) == str_length(ALT), "SNV", dplyr::if_else(str_length(REF) > str_length(ALT), "DEL", "INS")),
-         var_length = str_length(ALT) - str_length(REF),
-         var_class  = dplyr::if_else(var_length == 0, var_type, 
-                      dplyr::if_else(abs(var_length) == 1,  paste0(var_type, "1"),
-                      dplyr::if_else(abs(var_length) >  50, paste0(var_type, ">50"), paste0(var_type, "<50")))))
+## Index vcf files and merge them
+tabix input
+bcftools merge -m none -Oz `r config_ad$params_merge_vcf_files` -o output list_vcf_files
 
-variants.control <- variants %>% filter(Indiv %in% dplyr::filter(barcode_sheet, condition == "WT")$plant, gt_AR_corr > 0) 
-variants.freq_corr <- variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr)) %>% 
-  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(ID, target,amplicon) %>% count(name="freq")
+# Run tdnascan for vaector check
 
-variants <- variants %>% dplyr::filter(!ID %in% variants.control$ID) %>% dplyr::full_join(variants.freq_corr, by = c("ID", "target","amplicon"))
-#DT::datatable(variants)
-```
+## Cut multi-fasta vector file in individual files
+faSplit byname input output_directory `r config_ad$params_cut_vector`
 
-### nbr of variants per target and individu {.tabset .tabset-fade}
+## run tdnascan
+python2.7 tdnascan.py -1 input_fq1 -2 input_fq2 -t vector_sequence -g reference -p vector_name -i script_directory -d output_directory `r config_ad$params_tdnascan`
 
-```{r plot_nbr_variant ,results='asis',echo=TRUE}
-for (i in unique(bed.tab$target)){
-  cat("####",i,"\n")
-g<-variants  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr), target==i)  %>% 
-  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(Indiv, target,amplicon) %>% count(name="nbr_variants") %>%
-  ggplot() + geom_boxplot(aes(y=nbr_variants, x=amplicon))
-print(g)
-g2<- variants  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr), target==i)  %>%
-  select(ID, Indiv, target, var_type, amplicon) %>% unique() %>% group_by(Indiv, target, var_type,amplicon) %>% count(name="nbr_variants") %>%
-  ggplot() + geom_boxplot(aes(y=nbr_variants, x=var_type)) + facet_grid(~amplicon)
-print(g2)
-g3<- variants  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>%
-  select(ID, Indiv, target, var_class,amplicon) %>% unique() %>% group_by(Indiv, target, var_class,amplicon) %>% count(name="nbr_variants") %>%
-  ggplot() + geom_boxplot(aes(y=nbr_variants, x=target)) + facet_grid(amplicon~var_class) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylim(0, 10)
-print(g3)
-cat('\n\n')
-}
-
-```
-</div>
-
-### count per variant type and Indel length -Allele frequency in genotypes {.tabset .tabset-fade}
-
-```{r count_allele_by_type,results='asis',echo=TRUE}
-for (i in unique(bed.tab$target)){
-cat("####",i,"\n")
-g1<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>%  select(ID, Indiv, target) %>% unique() %>% group_by(ID, target) %>% count(name="freq") %>% ggplot() + 
-	geom_bar(aes(as.factor(freq), fill=target)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(g1)
-g2<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>% select(ID, Indiv, var_type) %>% unique() %>% group_by(ID, var_type) %>% count(name="freq") %>% ggplot() + geom_bar(aes(as.factor(freq), fill=var_type)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(g2)
-g3<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>% select(ID, Indiv, var_class) %>% unique() %>% group_by(ID, var_class) %>% count(name="freq") %>% ggplot() + geom_bar(aes(as.factor(freq), fill=var_class)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-print(g3)
-g4<-variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i) %>% select(ID, Indiv, var_class) %>% unique() %>% group_by(ID, var_class) %>% count(name="freq") %>% ggplot() + geom_bar(aes(as.factor(freq), fill=var_class)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-cat('\n\n')
-}
-```
-</div>
-
-## SV Calling
-
-```{r study_sv_calling }
-vcf_file<-paste0(workspace,"/05_sv_calling/variants.pbsv.annotate.vcf")
-sv.vcf <- vcfR::read.vcfR(vcf_file)
-
-vcfR.tid.sv.list <- vcfR::vcfR2tidy(sv.vcf)
-
-vcfR.tid.sv <- dplyr::full_join(vcfR.tid.sv.list$fix, vcfR.tid.sv.list$gt, by = c("ChromKey", "POS")) %>% 
-            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))  %>%
-            left_join(bed.tab, by=c("CHROM" = "rname"))%>% filter(POS >=start, POS <= stop) %>% select(-stop, -start) %>%
-            tidyr::separate(col = gt_AD, into = c("gt_ref_AD", "gt_alt_AD"), sep = ",", remove = FALSE) %>%
-            dplyr::mutate(gt_AR      = as.numeric(gt_alt_AD)/gt_DP,
-                          gt_AR_corr = dplyr::if_else(gt_AR < 0.2, 0, gt_AR),
-                          gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
-                          gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)))
-
-DT::datatable(vcfR.tid.sv)
-```
-
-### Number of variant per target {.tabset .tabset-fade}
-```{r plot_sv_calling ,results='asis',echo=TRUE} 
-for (i in unique(bed.tab$target)){
-cat("####",i,"\n")
-g1<-vcfR.tid.sv  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>% 
-  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(Indiv, target,amplicon) %>% count(name="nbr_variants") %>%
-  ggplot() + geom_boxplot(aes(y=nbr_variants, x=amplicon))+facet_grid(~target, scales = "free")
-print(g1)
-g2<-vcfR.tid.sv  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>% 
-  select(ID, target,amplicon, SVLEN) %>% unique() %>% group_by(target,amplicon, SVLEN) %>% count(name="sv_length") %>%
-  ggplot() + geom_boxplot(aes(y=sv_length, x=amplicon))+facet_grid(~target, scales = "free")
-#print(g2)
-g3<-vcfR.tid.sv  %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr),target==i)  %>%
-  select(ID, target, SVLEN,amplicon) %>% unique() %>% group_by(target, SVLEN,amplicon) %>%
-  ggplot() + geom_point(aes(y=SVLEN, x=amplicon))+facet_grid(~target, scales = "free")
-#print(g3)
-cat('\n\n')
-}
-```
-</div>
-
-# Source
-<a download="Report_snpeff_pbsv.html" href="`r base64enc::dataURI(file =params$rmd, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on pbsv data</a>
-<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>
-<a download="Report_snpeff.html" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on GenomicsVCF data</a>
+## run tdnascan annotate
+python2.7 tdnascan_annotate.py -i input -f annotation_file -o output `r config_ad$params_tdnascan_annotate`

### 2023-04-07 script/report.Rmd
merge modification

@@
 title: "MSGenova Rmakrdown Report"
-author: "Gabrièle Adam"
+author: " "
 date: "2022-08-16"

### 2023-04-09 script/ms_report.Rmd
add report; repove 06_report/

@@
 config_ad<-read_yaml(config_advanced_file)
-```
-
-```{r define_function}
-search_identical_insertion<-function(final_table,threashold){
-  list_index<-c()
-  # we are gonna filter by vector, then order table bu chr, position, and sample. Next compute the distance between 2 lines (0 then che change).Next use a threshold and filter this table
-  new_tab<-as_tibble(final_table) %>% dplyr::arrange(vector,Chr,Breakpoint,sample)
-  for (i in 1:nrow(new_tab)){
-    if(i ==1){
-      #case for line =1
-      next
-    }
-    if((new_tab[i,1] == new_tab[i-1,1]) & ((new_tab[i,2] - new_tab[i-1,2]) <= threashold )){
-      list_index <- append(list_index,c(i-1,i))
-    }
-  }
-  # now get line in final table
-  identical_insertion<-new_tab[unique(list_index),]
-  return(identical_insertion)
-}
-
-freq_length_bewteen_insertion<-function(final_table){
-  # Must be done by sample
-  length_between<-c()
-  for (v in unique(final_table$sample)){
-    for (spl in unique(final_table$sample)){
-  new_tab<-as_tibble(final_table)  %>% dplyr::filter(sample==spl & vector==v) %>% dplyr::arrange(Chr,Breakpoint)
-  if(nrow(new_tab) != 0){
-  for (i in 1:nrow(new_tab)){
-    if(i ==1){
-      #case for line =1
-      next
-    }
-   if((new_tab[i,1] == new_tab[i-1,1])){
-    length_between <-append(length_between,as.integer(new_tab[i,2] - new_tab[i-1,2]))
-   }
-  }
-  }
-    }
-  }
-  return(length_between)
-}
-
-read_file<-function(pathway){
-sample<-list.files(pathway)
-vector<-list.files(paste0(pathway,"/",sample[1]))
-i=0
-for (spl in sample){
-  for (v in vector){
-    table<-read.table(paste0(pathway,"/",spl,"/",v,"/5.",v,"_insertion.annotated.bed"),header = FALSE,col.names = c("Chr","Breakpoint","SuppRead","TDNA_info","Orientation","Freq","Annotation"))
-    if(! nrow(table)==0){
-      #add sample and vector columns
-      table["sample"]<-spl
-      table["vector"]<- v
-      if(i==0){
-        final_table<-table
-        i=i+1
-      }else{
-        final_table<-rbind(table,final_table)
-      }
-      i=i+1
-    }
-  }
-}
-if (exists("final_table")){
-return(final_table)}
-else{return(table)}
-}
-```
-
-`r ifelse(vector_calling,"# Study Plasmidic insertion","")`
-
-```{r define_variable}
-pathway=paste0(workspace,"/05_tdnascan")
-sample<-list.files(pathway)
-print(sample)
-vector<-list.files(paste0(pathway,"/",sample[1]))
-print(vector)
-# read sample file
-sample_sheet<-read.table(paste0(workspace,"/genome/sample_sheet"),sep="\t",header=TRUE)
-```
-
-```{r read_input}
-final_tab<-read_file(pathway)
-DT::datatable(final_tab)
-if(nrow(final_tab) >0){
-final_tab2<- final_tab %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector,sample)
-DT::datatable(final_tab2)
-presence_of_insertion<-TRUE
-}else{
-  presence_of_insertion<-FALSE
-}
-```
-
-`r ifelse(presence_of_insertion,"## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}","")`
-
-```{r barplot_snp, results='asis',eval=presence_of_insertion}
-#For one vector, look if it insert at the same position for several sample with a threshold for space between position
-for (v in vector){
-  sub_table<-filter(final_tab,vector==v) %>% group_by(Chr,sample) %>% dplyr::count(name="nbr_inster_by_chr")
-  #datatable(sub_table)
-  #count_tab<- count(sub_table,Chr)
-  if(nrow(sub_table)> 0){
-    cat("###",v,"\n")
-  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity",position=position_dodge())+ylab("Numbre of insertion by chromosome")+xlab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
-  print(g)
-  }
-  cat('\n\n')
-}
-```
-
-# Does several samples have the same insertions ?
-```{r same_insertions, eval=presence_of_insertion}
-essai<-search_identical_insertion(final_tab,5000)
-DT::datatable(essai)
-```
-
-`r ifelse(vector_calling,"# Barplot representation of the insertions in each sample","")`
-
-```{r plot_by_chr_tuto, eval=presence_of_insertion, echo = presence_of_insertion}
-#Establish list of sample without any insertion
-no_insertion<-c()
-#gff3<-"~/Documents/Msgenova/ref_arabido/Ath-Col0-tair10-WG/Arabidopsis_thaliana.TAIR10.41.gff3"
-# read gff3 and extract size of chromosome ?? - > Jump to final table
-##sequence-region   1 1 30427671
-##sequence-region   2 1 19698289
-##sequence-region   3 1 23459830
-##sequence-region   4 1 18585056
-##sequence-region   5 1 26975502
-##sequence-region   Mt 1 366924
-##sequence-region   Pt 1 154478
-table_size_chr<-data.frame(c(paste0("Chr",c(1:5,"Mt","Pt"))),c(30427671,19698289,23459830,18585056,26975502,366924,154478))
-colnames(table_size_chr)<-c("Chr","Size")
-for (spl in unique(final_tab$sample)){
-  sub_table<-dplyr::filter(final_tab,sample==spl)
-  if(nrow(sub_table)==0){
-    print("yes")
-    no_insertion<-append(no_insertion,spl)
-  }else{
- bp<- graphics::barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl),horiz=TRUE)
- rownames(bp)<-table_size_chr$Chr
-    with(sub_table,segments(sub_table$Breakpoint,bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),bp[sub_table$Chr,]+0.5,col=sub_table$vector)
-  #print(bp)
-  }
-}
-```
-
-```{r no_insertions, eval=presence_of_insertion}
-print("List of vector without any plasmid insertion")
-print(no_insertion)
-
 ```

### 2023-04-09 script/report.Rmd
add report; repove 06_report/

@@
 ---
-title: "MSGenova Rmakrdown Report"
+title: "to change"
 author: " "
@@
 ```{r setup, include=FALSE}
-workspace<-params$result_dir
-knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)
+# title: "`r params$title`"
+knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=TRUE, message=FALSE)
 ```
 
-```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
+```{r }
 library(tidyverse)
@@
 
-config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
-config_ad<-read_yaml(config_advanced_file)
 ```
+
+```{r}
+workspace<-params$result_dir
+```
+
+## Trimming parameters
+
+```{r trimmomatic}
+trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
+dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
+DT::datatable(dat)
+```
+
+## Mapping parameters
+
+### Flagstat results
+
+```{r flagstat}
+library(tidyverse)
+file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(file.exists(file)){
+    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
+      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
+                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
+                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
+                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
+    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")
+
+  }
+
+DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
+#count
+nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
+summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
+nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
+ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
+#ratio
+ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
+names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
+ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
+summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
+ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
+ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()
+
+```
+
+```{r SNV_InDels_calling.Rmd}
+
+  out <- NULL
+
+  out <- c(out, knitr::knit_child("SNV_InDels_calling.Rmd"), quiet=TRUE)
+
+knitr::asis_output(out)
+
+```
+
+```{r SV_calling.Rmd}
+
+  out <- NULL
+
+  out <- c(out, knitr::knit_child("SV_calling.Rmd"), quiet=TRUE)
+
+knitr::asis_output(out)
+
+```
+
+```{r tDNA_scan.Rmd}
+
+  out <- NULL
+
+  out <- c(out, knitr::knit_child("tDNA_scan.Rmd"), quiet=TRUE)
+
+knitr::asis_output(out)
+
+```
+
 # Tools used
 
-```{r tools_used}
-space=paste0(workspace,"/00_logs/envs")
-for (file in list.files(space)){
-  env<-read_yaml(paste0(space,"/",file))
-  for (envi in env$dependencies){
-    line=sub("=="," version ",envi)
-    cat(line,"\n")
-  }
-}
+```{r tools}
+
+  out <- NULL
+
+  out <- c(out, knitr::knit_child("tools.Rmd"), quiet=TRUE)
+
+knitr::asis_output(out)
+
 ```
-
-# Trimming
-trimmomatic PE `r config_ad$params_trimmomatic ` input_r1 input_r2 output_r1 output_r2 TRAILING:3 2> log
-
-# Alignments
-
-## alignments
-bwa mem -R "@RG\tID:sample\tSM:sample\tLB:Illumina\tPL:Illumina\tPU:sample" -M `r config_ad$params_bwa` reference input_fq1 input_fq2 > output
-
-## convert sam to bam
-picard SortSam --INPUT input --OUTPUT output --SORT_ORDER coordinate `r config_ad$params_sort_sam`
-
-## mark duplicates
-picard MarkDuplicates --INPUT input --OUTPUT{output -M metrics `r config_ad$params_mark_duplicate`
-
-# QC bam
-
-## flagstat
-samtools flagstat `r config_ad$params_flagstat ` input > output
-
-## samtools coverage
-samtools coverage -o output `r config_ad$params_samtools_coverage ` input
-
-## samtools depth
-samtools depth `r config_ad$params_samtools_depth`  input > output
-
-# SV Calling
-
-## Run haplotyCallze
-gatk HaplotypeCaller -R reference -I input -O output -L region --native-pair-hmm-threads 16 `r config_ad$params_haplotype_caller`
-
-## Run GenomicsDB
-gatk GenomicsDBImport variant --genomicsdb-workspace-path output --intervals {params.region} `r config_ad$params_genomics_db`
-
-## Run GenotypeGVCF
-gatk GenotypeGVCFs --reference reference -V gendb://input -O output `r config_ad$params_genotype_gvcf`
-
-## Merge gvcf
-picard GatherVcfs list_of_vcf_files O=output `r config_ad$params_gathervcf`
-
-## Build snpeff database
-snpEff build -c output `r config_ad$params_snpeff_build` -gff3 -v reference -dataDir .
-
-## Annotate variant
-snpEff eff -c input `r config_ad$params_snpeff_ann` -dataDir . reference -s output.html input > output
-
-# Run pindel
-## run pindel
-pindel `r config_ad$params_run_pindel`  -f reference -i input -o output
-
-## convert pindel to vcf
-pindel2vcf `r config_ad$params_convert_pindel` -r reference -P sample name -R xxx -d xx -v output
-
-## Index vcf files and merge them
-tabix input
-bcftools merge -m none -Oz `r config_ad$params_merge_vcf_files` -o output list_vcf_files
-
-# Run tdnascan for vaector check
-
-## Cut multi-fasta vector file in individual files
-faSplit byname input output_directory `r config_ad$params_cut_vector`
-
-## run tdnascan
-python2.7 tdnascan.py -1 input_fq1 -2 input_fq2 -t vector_sequence -g reference -p vector_name -i script_directory -d output_directory `r config_ad$params_tdnascan`
-
-## run tdnascan annotate
-python2.7 tdnascan_annotate.py -i input -f annotation_file -o output `r config_ad$params_tdnascan_annotate`

### 2023-04-09 script/tools.Rmd
add report; repove 06_report/

```{r}

config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
config_ad<-read_yaml(config_advanced_file)
```
# Tools used

```{r tools_used}
space=paste0(workspace,"/00_logs/envs")
for (file in list.files(space)){
  env<-read_yaml(paste0(space,"/",file))
  for (envi in env$dependencies){
    line=sub("=="," version ",envi)
    cat(line,"\n")
  }
}
```

# Trimming
trimmomatic PE `r config_ad$params_trimmomatic ` input_r1 input_r2 output_r1 output_r2 TRAILING:3 2> log

# Alignments

## alignments
bwa mem -R "@RG\tID:sample\tSM:sample\tLB:Illumina\tPL:Illumina\tPU:sample" -M `r config_ad$params_bwa` reference input_fq1 input_fq2 > output

## convert sam to bam
picard SortSam --INPUT input --OUTPUT output --SORT_ORDER coordinate `r config_ad$params_sort_sam`

## mark duplicates
picard MarkDuplicates --INPUT input --OUTPUT{output -M metrics `r config_ad$params_mark_duplicate`

# QC bam

## flagstat
samtools flagstat `r config_ad$params_flagstat ` input > output

## samtools coverage
samtools coverage -o output `r config_ad$params_samtools_coverage ` input

## samtools depth
samtools depth `r config_ad$params_samtools_depth`  input > output

# SV Calling

## Run haplotyCallze
gatk HaplotypeCaller -R reference -I input -O output -L region --native-pair-hmm-threads 16 `r config_ad$params_haplotype_caller`

## Run GenomicsDB
gatk GenomicsDBImport variant --genomicsdb-workspace-path output --intervals {params.region} `r config_ad$params_genomics_db`

## Run GenotypeGVCF
gatk GenotypeGVCFs --reference reference -V gendb://input -O output `r config_ad$params_genotype_gvcf`

## Merge gvcf
picard GatherVcfs list_of_vcf_files O=output `r config_ad$params_gathervcf`

## Build snpeff database
snpEff build -c output `r config_ad$params_snpeff_build` -gff3 -v reference -dataDir .

## Annotate variant
snpEff eff -c input `r config_ad$params_snpeff_ann` -dataDir . reference -s output.html input > output

# Run pindel
## run pindel
pindel `r config_ad$params_run_pindel`  -f reference -i input -o output

## convert pindel to vcf
pindel2vcf `r config_ad$params_convert_pindel` -r reference -P sample name -R xxx -d xx -v output

## Index vcf files and merge them
tabix input
bcftools merge -m none -Oz `r config_ad$params_merge_vcf_files` -o output list_vcf_files

# Run tdnascan for vaector check

## Cut multi-fasta vector file in individual files
faSplit byname input output_directory `r config_ad$params_cut_vector`

## run tdnascan
python2.7 tdnascan.py -1 input_fq1 -2 input_fq2 -t vector_sequence -g reference -p vector_name -i script_directory -d output_directory `r config_ad$params_tdnascan`

## run tdnascan annotate
python2.7 tdnascan_annotate.py -i input -f annotation_file -o output `r config_ad$params_tdnascan_annotate`

### 2023-04-11 script/quality_control.Rmd
add rmd for quality control

## Trimming parameters

```{r trimmomatic}
trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
DT::datatable(dat)
```

## HTML

## Mapping parameters

### Flagstat results

```{r }

file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
if(file.exists(file)){
    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

  }

DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')

#count
nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")

p <- ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=variable)) + geom_boxplot()  + 
  xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()

print(p)

#ratio
ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")

ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")

summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% 
  summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))

p <- ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=variable)) + 
  xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + 
  ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()

print(p)

```

### Coverage analysis
```{r coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
  }
})

cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
#cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
#cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min))

p <-ggplot(data = cov) + geom_boxplot(aes(y=meandepth, x=rname)) + 
  xlab("") + ggtitle("Plot coverage")  + 
  theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()

print(p)

```

### 2023-04-11 script/SNV_InDels_calling.Rmd
update rmd

# Small variants and Indels

```{r read vcf,message=FALSE, warning=FALSE, echo=FALSE}
# lire le ficher vcf
vcf <- read.vcfR(file = vcf.file, verbose = FALSE)
# transformer le vcfR en tibble
vcf.tbl.list <- vcfR2tidy(vcf) 
# concaténer fix et gt
# rajouter l'info ID
# extraire la valeur AD pour chaque allele
# calculer l'allele ratio
vcf.tbl <- dplyr::full_join(vcf.tbl.list$fix , vcf.tbl.list$gt, by = c("ChromKey", "POS")) %>%
           dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID)) %>% 
           dplyr::filter(!str_detect(ALT, ",")) %>%
           tidyr::separate(col = gt_AD, into = c("gt_ref.AD", "gt_alt.AD"), sep = ",", remove=FALSE) %>%
           dplyr::mutate(gt_ref.AD = as.numeric(gt_ref.AD), gt_alt.AD = as.numeric(gt_alt.AD)) %>%
           dplyr::mutate(gt_AR = if_else(gt_DP!=0 & !is.na(gt_DP), gt_alt.AD/gt_DP, as.numeric(NA))) %>%
           dplyr::mutate(TYPE = if_else(str_length(REF)>1 || str_length(ALT)>1, "INDEL", "SNV")) %>%
           dplyr::filter(!Indiv %in% c("SOJ-2-6-2-2-bis", "SOJ-5-3-3-1-bis", "SOJ-3-5-3-1"))

# la taille du tableau
#dim(vcf.tbl)
#head(vcf.tbl)

samples <- unique(vcf.tbl$Indiv)

samples
```

# contrôle qualité et nettoyage des données

## Les métriques de qualité

```{r, score quality, out.width="50%", fig.show="hold", message=FALSE, warning=FALSE, echo=FALSE}

# Score de qualité
p <- ggplot(data=vcf.tbl) + geom_density(aes(QUAL)) + 
          xlab("score de qualité") + ylab("densité") + xlim(c(0,200))
print(p)

# distribution de la couverture des positions variantes
p <- ggplot(data=vcf.tbl) + geom_boxplot(aes(y=gt_DP, x=Indiv, color=Indiv), outlier.shape = NA) + 
          ylab("DP of variants") + ylim(0, 100) +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
print(p)

# Distribution de l'allèle ratio
p <- ggplot(data=vcf.tbl) + geom_density(aes(gt_AR, color=Indiv)) + 
          xlab("Ratio de l'allèle alternatif") + ylab("densité")

if (length(samples) > 15){
  p <- p + theme(legend.position="none")
}

print(p)

# compter le nombre de positions non couvertes par échantillon.
NA_count_samples <- dplyr::select(vcf.tbl, ID, Indiv, gt_DP) %>% 
                    dplyr::group_by(Indiv) %>% dplyr::count(is.na(gt_DP)) %>%
                    dplyr::filter(`is.na(gt_DP)` == TRUE)

p <- ggplot(data=NA_count_samples) +  geom_bar(aes(y=n, x=Indiv), stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("") + 
  ylab("Positions manquées (non couvertes)")
print(p)

```

## Filtres qualité
* score de qualité : > `r qual.min`
* nombre de reads qui couvrent la position variant : > `r DP.min`
* nombre de reads qui portent l'allèle alternatif : > 1
* retirer les positions avec au moins un échantillon non couvert.

```{r filtre qual, message=FALSE, warning=FALSE, echo=FALSE}

 vcf.tbl.q <- dplyr::filter(vcf.tbl, QUAL >= qual.min) %>%
              dplyr::mutate(gt_alt.AD = if_else(gt_alt.AD < 2, 0, as.numeric(gt_alt.AD)),
                            gt_DP     = if_else(gt_DP < DP.min, as.numeric(NA), as.numeric(gt_DP))) %>%
              dplyr::mutate(gt_AR = if_else(gt_DP!=0 & !is.na(gt_DP), gt_alt.AD/gt_DP, as.numeric(NA)))

#dim(vcf.tbl.q)

# vcf.tbl.count <- vcf.tbl %>% select(ID, gt_AR, Indiv) %>%  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
#   dplyr::mutate(NA_count = rowSums(is.na(dplyr::select(., -ID)))) %>% 
#   dplyr::mutate(count_var = rowSums(0 != (dplyr::select(., -ID, -NA_count)))) %>% 
#   select(ID, NA_count, count_var)

vcf.tbl.q.count <- vcf.tbl.q %>% select(ID, CHROM, gt_AR, Indiv) %>%  
  reshape2::dcast(CHROM+ID~Indiv, value.var = "gt_AR") %>%
  dplyr::mutate(NA_count.q = rowSums(is.na(dplyr::select(., -ID)))) %>% 
  dplyr::mutate(count_var.q = rowSums(0 != (dplyr::select(., -ID, -NA_count.q)))) %>% 
  dplyr::filter(NA_count.q == 0 , !is.na(count_var.q))

vcf.tbl.q <- vcf.tbl.q %>% filter(ID %in% vcf.tbl.q.count$ID)

#full_join(vcf.tbl.q.count , vcf.tbl.count , by="ID") %>% filter(NA_count != NA_count.q)

```

# Statistiques descriptives

## Nombre des variants par échantillon.

```{r, out.width="90%", message=FALSE, warning=FALSE, echo=FALSE}
vcf.tbl.q %>% dplyr::filter(gt_AR > 0) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
  unique() %>% dplyr::group_by(Indiv) %>% dplyr::count(name="nbr_var") %>% 
  ggplot() + geom_bar(aes(x=Indiv, y=nbr_var, fill=Indiv), stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") + 
  xlab("") + ylab("# variants")
```

## Frequences alléliques dans les échantillons.

```{r, out.width="60%", message=FALSE, warning=FALSE, echo=FALSE}
# retirer les lignes variant-echantillon avec un allèle ref. 
vcf.tbl.q.var <- dplyr::filter(vcf.tbl.q, gt_AR > 0)

# Découper la la ligne et dire ce que chacune fait?
ID.uniq <- vcf.tbl.q.var %>% dplyr::select(ID, Indiv) %>% dplyr::group_by(ID) %>% count()

```

```{r, out.width="90%", message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
df <- dplyr::full_join(ID.uniq, vcf.tbl.q.var, by="ID") %>% 
      dplyr::select(Indiv, n) %>% dplyr::group_by(n, Indiv) %>% dplyr::count()

p <- ggplot(data=df, aes(x=as.factor(n), y=nn)) + geom_boxplot(outlier.shape = NA) +
  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")

if (length(samples) > 15){

  p <- p + theme(legend.position="none")
}
print(p)
```

# recherche de pics

* points gris : variations ref ou fréquentes (communes à 5 échantillons et plus).
* points verts : variations non ref et non fréqunetes.
* points rouges : variations non ref et non fréquentes avec un ratio égales ou sup à `r hom`.

```{r, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
vcf.tbl.q.var <- dplyr::filter(vcf.tbl.q, gt_AR > 0, !is.na(gt_AR))

ID.ref  <- dplyr::filter(vcf.tbl.q.var, Indiv == ref) 
#vcf.tbl.non.ref <- dplyr::filter(vcf.tbl.q.var, ! ID %in% ID.ref$ID)

#dim(vcf.tbl.non.ref)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}

# test sans dplyr::select -> error pour illuster pourquoi il est important de rajouter le nom du package avnat la fonction
#ID.uniq <- select(vcf.tbl.non.ref, ID, Indiv) %>% group_by(ID) %>% count() %>% filter(n < 2) 
ID.uniq <- dplyr::select(vcf.tbl.q.var, ID, Indiv) %>% group_by(ID) %>% count() %>% filter(n < 2)

vcf.tbl.uniq <- dplyr::filter(vcf.tbl.q.var, ID %in% ID.uniq$ID)
vcf.tbl.nonref  <- dplyr::filter(vcf.tbl.q.var, !ID %in% ID.ref$ID, ID %in% filter(vcf.tbl.q.count, count_var.q < 5)$ID)
#
#dim(vcf.tbl.uniq)

```

```{r plot all indiv, warning=FALSE, message=FALSE, fig.height=2, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}

mutations.list <- list()

# une boucle sur l'ensemble des échantillons
for(ind in unique(vcf.tbl$Indiv)){

  print(ind)

#  vcf.tbl.ind         <- dplyr::filter(vcf.tbl.q,       Indiv==ind)
  vcf.tbl.q.var.ind   <- dplyr::filter(vcf.tbl.q.var,   Indiv==ind)
  #vcf.tbl.uniq.ind    <- dplyr::filter(vcf.tbl.uniq,    Indiv==ind)
  vcf.tbl.uniq.ind  <- dplyr::filter(vcf.tbl.nonref,     Indiv==ind)

  hom.chr <- dplyr::filter(vcf.tbl.uniq.ind, gt_AR > hom) %>% dplyr::select(CHROM) %>% unique()
  var.to.smooth <- dplyr::filter(vcf.tbl.uniq.ind, CHROM %in% hom.chr$CHROM, gt_AR >= AR.min)

  # stocker le res dans une list
  mutations.list[[ind]] <- var.to.smooth %>% dplyr::filter(gt_AR > hom)

  p <- ggplot() + ylim(0,1) +
      geom_point( data=vcf.tbl.q.var.ind,    aes(x=POS, y=gt_AR), size=0.2, color="grey80") + 
      geom_point( data=vcf.tbl.uniq.ind,     aes(x=POS, y=gt_AR), size=0.3, color="green") + 
      geom_point( data=mutations.list[[ind]],aes(x=POS, y=gt_AR), size=0.5, color="red") + 
      geom_smooth(data=var.to.smooth,        aes(x=POS, y=gt_AR), size=0.5) +
      facet_grid( ~ CHROM, scale="free",  space = "free") +
      theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0),
            axis.text.y = element_text(size = rel(0.8), hjust = 0)) +
      scale_x_continuous(breaks = seq(0, max(vcf.tbl$POS), 1e+07))
  print(p)

}

# concatener les tibbles
mutations <- purrr::reduce(mutations.list, rbind) %>% select(CHROM, ID, QUAL, gt_DP, gt_ref.AD, gt_alt.AD, gt_AR, Indiv, ANN) %>% separate(ANN, c("tmp1","eff", "tmp2", "gene"), "\\|") %>% select(-paste0("tmp",1:2))

mutations %>% select(Indiv, ID, eff) %>% filter(!eff %in% c("downstream_gene_variant", "intergenic_region", "upstream_gene_variant")) %>% unique() %>% group_by(Indiv, eff) %>% count(name = "count")

write.table(mutations, file="SOJ_filt2_mutation_Ho.xls", quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t")

```

### 2023-04-11 script/report.Rmd
update rmd

@@
 ---
-title: "to change"
+title: "Report"
 author: " "
@@
 # title: "`r params$title`"
+workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
+#workspace<-params$result_dir
+
+sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+
 knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=TRUE, message=FALSE)
@@
 ```{r }
-library(tidyverse)
 library(DT)
 library(ggplot2)
-library(plyr)
+library(vcfR)
 library(dplyr)
@@
 ```{r}
-workspace<-params$result_dir
-```
 
-## Trimming parameters
-
-```{r trimmomatic}
-trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
-dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
-DT::datatable(dat)
-```
-
-## Mapping parameters
-
-### Flagstat results
-
-```{r flagstat}
-library(tidyverse)
-file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
-if(file.exists(file)){
-    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
-      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
-                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
-                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
-                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
-    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")
-
-  }
-
-DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
-#count
-nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
-summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
-nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
-ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
-#ratio
-ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
-names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
-ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
-summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
-ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
-ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()
+sample_sheet <- read.table(file = sample_sheet.file, header = TRUE, sep = "\t")
 
@@
 
-```{r SNV_InDels_calling.Rmd}
+```{r quality_control.Rmd}
+
+  out <- NULL
+
+  out <- c(out, knitr::knit_child("quality_control.Rmd"), quiet=TRUE)
+
+knitr::asis_output(out)
+
+```
+
+```{r SNV_InDels_calling.Rmd, eval=FALSE}
+
+vcf.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/03_snv_indels_calling/SOJ_snv_indel.vcf"
 
@@
 
-```{r SV_calling.Rmd}
+```{r SV_calling.Rmd, eval=FALSE}
 
@@
 
-```{r tDNA_scan.Rmd}
+```{r tDNA_scan.Rmd, eval=FALSE}
 
@@
 
-# Tools used
-
-```{r tools}
+```{r tools, eval=FALSE}
 

### 2023-04-14 script/per_sample_ins_desc.Rmd
Add tdnascan report.

### `r sample` {.tabset}

#### Summary of insertion reads

```{r insertion, echo=FALSE}

insRead.list <- list()

for(vector_i in perSample_vectorList[[sample]]$both.list){

  insertion_read_file <- paste0(path_results,"/results/05_tdnascan","/",sample,"/",vector_i,"/4.insertionRead.txt")

  if(file.exists(insertion_read_file) & file.size(insertion_read_file) != 0 ){

    insRead.list[[paste0(sample,"_",vector_i)]] <- read.table(insertion_read_file,sep="\t") %>% 
      dplyr::mutate(sample = sample) %>% 
      dplyr::mutate(vector = vector_i)

  }
}

insRead <- purrr::reduce(insRead.list, rbind)

insRead.summary <- insRead %>% 
dplyr::mutate(insRead,V1=str_remove_all(V1,pattern="[\\(\\)]")) %>%   tidyr::separate(V1,into=c("chromosome","position","type","tdna_border","tdna_position","orientation"),sep=",") %>% dplyr::mutate(position=as.numeric(position)) %>% dplyr::group_by(chromosome,position,type,tdna_border,tdna_position,orientation,vector) %>%
dplyr::count()

DT::datatable(insRead.summary)
```

```{r depthNGS, eval=TRUE, echo=FALSE}

depth <- read.table(paste0(path_results,"/results/02_mapping/depth/",sample,".depth"),h=TRUE) 
colnames(depth) <- c("rname", "position", "depth")
SIZE=10000

p <- list()

for(vector_j in unique(ins$vector)){

  ins.sample <- dplyr::filter(ins, sample == !!sample, vector == !!vector_j) %>% 
    dplyr::group_by(vector, insertion_id)

  depth.filt.sample <- list()

  if(dim(ins.sample)[1] != 0){

    for(i in 1:dim(ins.sample)[1]){

      depth.filt <- dplyr::filter(depth, 
                                  rname == ins.sample[i,]$Chromosome, 
                                  position >= (ins.sample[i,]$Position-SIZE),
                                  position <= (ins.sample[i,]$Position+SIZE))

      depth.filt <- dplyr::mutate(depth.filt, 
      insertion_id = paste0(sample,"_",ins.sample[i,]$vector,"_",rname,"_ins_",i), vector=ins.sample[i,]$vector)

      depth.filt.sample[[i]] <- depth.filt[seq(from=1,to=dim(depth.filt)[1],by=20),]

    }
    depth.plot <- purrr::reduce(depth.filt.sample, rbind)

    p[[vector_j]] <- ggplot(depth.plot) + geom_bar(aes(x=position, y=depth),stat="identity",colour="blue") + 
      theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
      ylab("Depth of coverage") + facet_wrap(vars(insertion_id),scales = "free", ncol=1) + 
      geom_vline(data=ins.sample, aes(xintercept = Position ,col="red")) + 
      geom_text(data=ins.sample,aes(x=Position,y=4,label=paste0("ins_pos: ",Position),colour="red",fontface = "bold"))
    #print(p[[vector_j]])
  }
}

```

#### Depth along insertion sites +- 10 kb {.tabset}

```{r, eval=TRUE, echo=FALSE, results='asis', fig.height=15}
for(vector_j in unique(ins$vector)){
  cat("##### Vector", vector_j, "\n")
  print(p[[vector_j]])
  cat("\n\n")
}
```

### 2023-04-14 script/per_sample_sv_desc.Rmd
Add tdnascan report.

### `r sample` {.tabset}

#### summary of homozygous SV

```{r sv_spec_plot_hom, echo=FALSE}

bam <- paste0(sample,".bam")

SV.NGS.hom <-  SV %>%  dplyr::filter(
    .data[[bam]]  %in% c("1/1"),SPEC == 1) %>% dplyr::arrange(.,CHROM,POS)

if(dim(SV.NGS.hom)[1] >=1){
SV.NGS.hom  %>% group_by(CHROM,SVTYPE) %>% count() %>% ggplot() + geom_bar(aes(x=CHROM,y=n,fill=SVTYPE),stat="identity") + scale_fill_manual(values=c("lightsalmon","royalblue1","green","magenta1","yellow"),drop=FALSE) + xlab("Chromosoms")
}

```

#### table of homozygous SV

```{r sv_spec_plot_table_hom, echo=FALSE}
DT::datatable(SV.NGS.hom %>% dplyr::select(.,-contains(".bam")))
```

#### summary of heterozygous SV

```{r sv_spec_plot_het, echo=FALSE}
bam <- paste0(sample,".bam")

SV.NGS.het <-  SV %>%  dplyr::filter(
    .data[[bam]]  %in% c("0/1"),SPEC == 1) %>% dplyr::arrange(.,CHROM,POS)

if(dim(SV.NGS.het)[1] >=1){
SV.NGS.het %>% group_by(CHROM,SVTYPE) %>% count() %>% ggplot() + geom_bar(aes(x=CHROM,y=n,fill=SVTYPE),stat="identity") + scale_fill_manual(values=c("lightsalmon","royalblue1","green","magenta1","yellow"),drop=FALSE)
}
```

#### table of heterozygous SV

```{r sv_spec_plot_table_het, echo=FALSE}
DT::datatable(SV.NGS.het %>% dplyr::select(.,-contains(".bam")))
```

### 2023-04-14 script/per_sample_vector_depth.Rmd
Add tdnascan report.

### `r sample`

```{r depthVec, eval=TRUE, echo=FALSE}

file <- paste0(path_results,"/results/02_mapping/depth/",sample,".depth")

if(file.exists(file)){
  depth <- read.table(file,h=TRUE) 
}

colnames(depth) <- c("rname", "position", "depth")

depth.vector <- dplyr::filter(depth, rname %in%  perSample_vectorList[[sample]]$both.list)

if(dim(depth.vector)[1] > 0){

ggplot(depth.vector) + geom_bar(aes(x=position, y=depth),stat="identity",colour="blue") + 
  theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ylab("Depth of coverage") + facet_wrap(vars(rname),scales = "fixed",ncol=1)
}

```

### 2023-04-14 script/tDNA_scan.Rmd
Add tdnascan report.

# parameters

```{r param}
path_results <- workspace
samplesheet <- sample_sheet
vector.sep <- ","
nb_common_samples=10
SVsize_CutOff = 500
```

# test sample sheet

```{r testSS, eval=TRUE, echo=FALSE}

# test de l'existence de sample_sheet:
# !! des colonnes additionnelles sont attendues pour le sample_sheet !!
# Si celles-ci sont complétées au démarrage du pipeline, cela posera
# problème avec le pipeline snakemake
# tdna (required): 1 to many separate by ","
# vector (optional): NA to many separate by ","

if(dir.exists(path_results) & file.exists(samplesheet)){
  sample_sheet <- read.table(samplesheet,sep="\t",header=TRUE)
}

# construction of the list of searched tdna and/or vector

perSample_vectorList <- as.list(1:dim(sample_sheet)[1])

for(k in 1:dim(sample_sheet)[1]){
  tmp.l <- list()
  tmp.l[["tdna.list"]] <- stringr::str_split(sample_sheet[k,"tdna"], pattern = vector.sep)[[1]]
  tmp.l[["vector.list"]] <- stringr::str_split(sample_sheet[k,"vector"],pattern = vector.sep)[[1]]
  tmp.l[["both.list"]] <- na.omit(c(tmp.l[["tdna.list"]],tmp.l[["vector.list"]]))
  perSample_vectorList[[k]] <- tmp.l
}
names(perSample_vectorList) <- sample_sheet$sample

```

# Load libraries

```{r load.lib, echo=FALSE,message=FALSE}
library(tidyverse)
library(vcfR)
```

# tDNA coverage statistics in comparison with other chromosoms:

This statistics will indicate the number of times the tDNA may be inserted and if it
is completely inserted.

## Mean depth:

```{r depth, echo=FALSE, warning=FALSE}

# reads the statistics of chromosoms and vectors mean depth

cov.list <- lapply(sample_sheet$sample, function(x){

  file <- paste0(path_results,"/results/02_mapping/coverage","/", x, ".coverage")

  if(file.exists(file)){
    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
  }
})

if(length(cov.list) != 0){

  cov <- purrr::reduce(cov.list, rbind)

  colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	
                     "covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample")

  # Mean depth (log10)
  g1 <- ggplot(cov) + geom_bar(aes(y=meandepth, x=rname),stat = "identity") + 
    xlab("") + ylab("Mean Depth") + ggtitle("Mean depth (log10)") + facet_wrap( ~ sample) +
    theme(axis.text.x=element_text(angle=90,hjust=1)) + theme_bw() + scale_y_continuous(trans='log10',limits=c(1,NA)) +
    coord_flip() + geom_text(aes(x = rname,y=10,label=paste0(round(meandepth,0)," x")),color="red")
  g1
}
```

## Chromosoms coverage (%):

```{r coverage, echo=FALSE}
if(length(cov) != 0){
  g2<-ggplot(cov) + geom_bar(aes(y=coverage, x=rname),stat = "identity") + xlab("") + ylab("% coverage") + ggtitle("Chromosoms coverage (%)")+facet_wrap( ~ sample)+theme(axis.text.x=element_text(angle=90,hjust=1)) + theme_bw() + coord_flip()
  g2
}
```

## Depth of coverage along vectors:

<!-- ### Graph per sample -->

```{r , message=FALSE, warning=FALSE, echo=FALSE}

  out <- NULL

  for (sample  in  unique(sample_sheet$sample)) {
      out <- c(out, knitr::knit_child("per_sample_vector_depth.Rmd", quiet=TRUE))  
    }

knitr::asis_output(out)
```

# TDNAscan results:

tdnascan was used to detect tdna insertion [@sun_tdnascan_2019].

## Table of detected insertion: {.tabset}

* Results description:
  + Chromosome: Chromosome number
  + Position: Start position of insertion in the reference genome
  + \#soft.clipped: The soft-clipped reads (CLR) are reads where one partial read of a single read perfectly mapped to the plant reference genome and the other partial read of the same single read perfectly mapped to the inserted T-DNA. 
  + \#Discordant: In tdnascan, the discordant reads (DIR) are defined as one read of a pair successfully mapped to the plant reference genome and the other read of the same pair mapped to part of the inserted T-DNA.
  + tdna.start: Start position of T-DNA sequence
  + tdna.end: End position of T-DNA sequence
  + orientation: Forward or reverse T-DNA inserted to reference genome
  + insertion zygotie: #soft.clipped/(\#soft.clipped + \#span) with \#span = \# reads MAPQ > 30 which span the Insertion site with at least 5bp at each side
  + annotation:
  + sample: 
  + insertion_id:

### Overview of exact insertion's positions

```{r bed.tab.plot, echo =FALSE,message=FALSE}

# Concatenate all tdnascan results

ins.list <- list()

for(x in sample_sheet$sample){

  for(vector_i in perSample_vectorList[[x]]$both.list){

    file <- list.files(path=paste0(path_results,"/results/05_tdnascan","/",x,"/",vector_i,"/"), 
                       pattern="annotated.bed", full.names = TRUE)

    if(length(file) != 0){
      if(file.size(file) != 0){

        tmp <- read.table(file, sep="\t",h=FALSE) %>% dplyr::mutate(sample = x, vector = vector_i)

        if(dim(tmp)[1] != 0){

          ins.list[[paste0(x,"_",vector_i)]] <- 
            dplyr::mutate(tmp, insertion_id = paste0(x,"_",vector_i,"_",V1,"_ins_",1:dim(tmp)[1]))

        }
      }
    }
  }
}

ins <- purrr::reduce(ins.list, rbind) 
ins <- separate(ins,V3,into=c("CLR","DIR"),sep=",")
ins <- separate(ins,V4,into=c("tdna_st","tdna_end"),sep=",")
ins <- mutate(ins,CLR = str_extract(CLR,pattern="(\\d)+"),
              DIR=str_extract(DIR,pattern="(\\d)+"),
              tdna_st=str_extract(tdna_st,pattern="(\\d)+"),
              tdna_end=str_extract(tdna_end,pattern="(\\d)+"))

names(ins)<-c("Chromosome","Position","#soft.clipped","#Discordant",
              "tdna.start","tdna.end","orientation","insertion zygotie",
              "annotation","sample","vector","insertion_id")

Chrom <- dplyr::filter(cov,rname %in% ins$Chromosome) %>% dplyr::select(.,c("rname","startpos","endpos")) 
Chrom <- dplyr::distinct(Chrom) %>% dplyr::rename(.,Chromosome=rname)

 ggplot(data=ins)+
     geom_rect(data=Chrom, mapping = aes(xmin=startpos,xmax=endpos,ymin=0.2,ymax=0.3),color="grey")+
     facet_grid(sample~Chromosome,space="free_x",scales="free_x") + 
   geom_jitter(data=ins, aes(x=Position,y=1,color=vector),size=2,shape=25) + ylim(c(0,2)) + 
    theme(axis.text.y=element_blank(),axis.ticks.y =element_blank(),axis.text.x= element_text(angle = 90), axis.title.y=element_blank(),axis.title.x=element_blank(),strip.background = element_blank(),panel.grid.minor = element_blank(),strip.text.y = element_text(angle=-90),panel.grid.major = element_blank(),strip.placement = "outside")

```

### Exact insertion's positions table

```{r bed.tab.view, echo =FALSE,message=FALSE}
DT::datatable(ins)
```

### Overview of approximativ breakpoint's positions

Overview of detected insertion before filtering for nearby breakpoints:

```{r nofiltre_view, echo = FALSE, message = FALSE}

ins.list.raw <- list()

for(x in sample_sheet$sample){

  for(vector_i in perSample_vectorList[[x]]$both.list){

    file <- list.files(paste0(path_results,"/results/05_tdnascan","/",x,"/",vector_i,"/"),
                       pattern="insertion.bed", full.names = TRUE)

    if(length(file) != 0){

      ins.list.raw[[paste0(x,"_",vector_i)]] <- read.table(file, sep="\t",h=TRUE) %>% 
        dplyr::mutate(sample = x) %>% dplyr::mutate(vector = vector_i) 
    }
  }
}

ins_raw <- purrr::reduce(ins.list.raw, rbind)

if(length(ins_raw) > 0){

  ins_raw <- tidyr::separate(ins_raw,SuppRead,into=c("CLR","DIR"),sep=",")
  ins_raw <- tidyr::separate(ins_raw,TDNA_info,into=c("tdna_st","tdna_end"),sep=",")
  ins_raw <- dplyr::mutate(ins_raw,CLR=str_extract(CLR,pattern="(\\d)+"),
                           DIR=str_extract(DIR,pattern="(\\d)+"),
                           tdna_st=str_extract(tdna_st,pattern="(\\d)+"),
                           tdna_end=str_extract(tdna_end,pattern="(\\d)+"))
  ins_raw <- dplyr::filter(ins_raw,str_detect(Breakpoint,"~")==TRUE)
  ins_raw <- dplyr::mutate(ins_raw,Breakpoint=as.numeric(str_remove(Breakpoint,"~")))

  names(ins_raw)<-c("Chromosome","Breakpoint","#soft.clipped","#Discordant",
                    "tdna.start","tdna.end","orientation","insertion zygotie",
                    "sample","vector")

  Chrom <- dplyr::filter(cov,rname %in% ins_raw$Chromosome) %>% dplyr::select(.,c("rname","startpos","endpos")) 
  Chrom <- dplyr::distinct(Chrom) %>% dplyr::rename(.,Chromosome=rname)

  ggplot(data=ins_raw)+
    geom_rect(data=Chrom, mapping = aes(xmin=startpos,xmax=endpos,ymin=0.2,ymax=0.3),color="grey")+
    facet_grid(sample~Chromosome,space="free_x",scales="free_x") + 
    geom_jitter(data=ins_raw, aes(x=Breakpoint,y=1,color=vector),size=2,shape=25) + ylim(c(0,2)) + 
    theme(axis.text.y=element_blank(),axis.ticks.y =element_blank(),axis.text.x= element_text(angle = 90), axis.title.y=element_blank(),axis.title.x=element_blank(),strip.background = element_blank(),panel.grid.minor = element_blank(),strip.text.y = element_text(angle=-90),panel.grid.major = element_blank(),strip.placement = "outside", ) 
}

```

### Approximativ breakpoints table

Table of detected insertion before filtering for nearby breakpoints:

```{r nofiltre_table, echo =FALSE,message=FALSE}
DT::datatable(ins_raw)
```

## Informativs reads description:

* This table describes informatives reads. The reads which mapped to part of the tDNA and which cover the insertion sites. 
  + for CLR: The reads are retained if they satisfied the criterion of the difference in the number of soft clipped bases in the TDNA and in the genome (<5). See the figure below.     
  + for DIR: The reads are retained if one mate match completly to the tDNA and the other to the genome.
  + TSP: tdna start border 
  + TEP: tdna end boder

<!-- ### Results per sample -->

```{r , message=FALSE, warning=FALSE, echo=FALSE}

  out <- NULL

  for (sample  in  unique(sample_sheet$sample)) {
      out <- c(out, knitr::knit_child("per_sample_ins_desc.Rmd", quiet=TRUE))

    }

knitr::asis_output(out)
```

# Do the tdna-insertion lines have specifics Structural Variation ?

* pindel [ref] was used to detect the following structural variations.
  + INV: Inversion (min_length=50): 
  + DUP:TANDEM: tandem duplication
  + INS: Insertion: Domestic insertion (DINS) of Foreign Insertion (FINS) or Translocation (TRA)
  + DEL: Deletion
  + RPL: Replacement: describes an insertion event around the breakpoint of a deletion event, which could in fact capture DUP, INV, DINS, FINS and TRA, thus occasionally resulting in duplicate SV calls with different assigned SV types.

* NB: 
  + pindel as a poor sensitivity to detect DINS or TRANS. It may not help to confirm tDNAscan results but may help to detect 
other SVs that might be induced by the tDNA lines transformation.
We will include other sv caller results such as delly to cross tdnascan results and to ensure of the lines background.
  + SVs were classified according to their genotypes: 1/1 = homozygous for the alternate allele and 1/0 = heterozygous
  They were not confirm thanks to igv, or by filtering allelic depth. 
  + A filter was applied to keep SV longer than 500.
  + A column NEXT_SV was add to indicate the distance (in base) to the next SV: it can be a common or specific one. 
  It will help to filter sv classified in specific sv whereas they are common to samples.

```{r sv,eval=TRUE, message=FALSE,echo=FALSE}

pindel <- paste0(path_results,"/results/04_pindel/",path_results,"_sv.vcf.gz")

# import pindel vcf

if(file.exists(pindel)){
  vcfR.inst <- vcfR::read.vcfR(pindel,verbose = FALSE)

  nsample <- dim(vcfR.inst@gt)[2]-1

  # transform to tidy
  vcfR.tid.list <- suppressMessages(vcfR::vcfR2tidy(vcfR.inst))

  # filter SV by SVsize_CutOff 
  sv <- vcfR.tid.list[[1]] %>% dplyr::mutate(.,ID=c(1:dim(vcfR.tid.list[[1]])[1]))
  sv.filt <- sv[which(abs(sv$SVLEN) > SVsize_CutOff),]

  if(dim(sv.filt)[1] > 0){
    # filter geno by ID in sv and by GT: 1/1 or 0/1
    geno <- vcfR.tid.list[[2]]  %>%  dplyr::mutate(.,ID=rep(c(1:dim(vcfR.tid.list[[1]])[1]),times=nsample))
    geno.filt <-  dplyr::filter(geno,ID %in% sv.filt$ID, gt_GT %in% c("1/1","0/1")) %>% dplyr::arrange(ID)
  }

  if(dim(geno.filt)[1] > 0){
    geno_tmp <- geno.filt %>% dplyr::select(ID,gt_GT,Indiv,ID) %>% tidyr::pivot_wider(names_from=Indiv,values_from=gt_GT)

    SV <- dplyr::left_join(geno_tmp ,sv.filt,by="ID")  %>% dplyr::arrange(ID) %>% dplyr::mutate(.,SVTYPE=as.factor(SVTYPE)) %>%  dplyr::select(.,ID,CHROM,POS,END,SVLEN,SVTYPE,contains(".bam")) %>% dplyr::mutate(.,DIST_NEXTSV=c(diff(POS),NA))

    tmp <- dplyr::select(SV,contains(".bam"))
    SV$SPEC <- rowSums(apply(! is.na(tmp),2,as.numeric))
  }
}
```

## Overview of SV repartition on chromosoms:

```{r SV_overview, echo=FALSE}
Chrom <- dplyr::filter(cov,rname %in% SV$CHROM) %>% dplyr::select(.,c("rname","startpos","endpos")) 
Chrom <- dplyr::distinct(Chrom) %>% dplyr::rename(.,CHROM=rname)

 ggplot(data=SV)+
     geom_rect(data=Chrom, mapping = aes(xmin=startpos,xmax=endpos,ymin=0,ymax=0.1),color="grey")+
     facet_grid(CHROM ~.,space="free_x",scales="free_x") + 
   geom_jitter(data=SV, aes(x=POS,y=1,color=as.factor(SPEC)),size=2,shape=25) + ylim(c(0,2)) + xlab("Positions") + 
    theme(axis.text.y=element_blank(),axis.ticks.y =element_blank(),axis.text.x= element_text(angle = 90), axis.title.y=element_blank(),axis.title.x=element_blank(),strip.background = element_blank(),panel.grid.minor = element_blank(),strip.text.y = element_text(angle=-90),panel.grid.major = element_blank(),strip.placement = "outside")+
          scale_colour_discrete(name  = "Number of \n samples per SV")

```

## Common SV {.tabset}

### summary of common SV

```{r commonsv_sum,eval=TRUE, message=FALSE,echo=FALSE}
sv_common <- dplyr::filter(SV, SPEC !=1 )
sv_common %>% group_by(CHROM,SVTYPE) %>% count() %>% ggplot() + geom_bar(aes(x=CHROM,y=n,fill=SVTYPE),stat="identity") + scale_fill_manual(values=c("lightsalmon","royalblue1","green","magenta1","yellow"),drop=FALSE) + xlab("Chromosoms")
```

### table

```{r commonsv_tab,eval=TRUE, message=FALSE,echo=FALSE}
DT::datatable(sv_common)
```

## Specific SV per sample:

<!-- ### Results per sample -->

```{r , message=FALSE, warning=FALSE, echo=FALSE}

  out <- NULL

  for (sample  in  unique(sample_sheet$sample)) {
      out <- c(out, knitr::knit_child("per_sample_sv_desc.Rmd", quiet=TRUE))

    }

knitr::asis_output(out)
```

### 2023-06-20 script/SNV_InDels_calling.Rmd
update report

@@
-# Small variants and Indels
+# SNV and small Indels
 
-```{r read vcf,message=FALSE, warning=FALSE, echo=FALSE}
-# lire le ficher vcf
-vcf <- read.vcfR(file = vcf.file, verbose = FALSE)
-# transformer le vcfR en tibble
-vcf.tbl.list <- vcfR2tidy(vcf) 
-# concaténer fix et gt
-# rajouter l'info ID
-# extraire la valeur AD pour chaque allele
-# calculer l'allele ratio
-vcf.tbl <- dplyr::full_join(vcf.tbl.list$fix , vcf.tbl.list$gt, by = c("ChromKey", "POS")) %>%
-           dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID)) %>% 
-           dplyr::filter(!str_detect(ALT, ",")) %>%
-           tidyr::separate(col = gt_AD, into = c("gt_ref.AD", "gt_alt.AD"), sep = ",", remove=FALSE) %>%
-           dplyr::mutate(gt_ref.AD = as.numeric(gt_ref.AD), gt_alt.AD = as.numeric(gt_alt.AD)) %>%
-           dplyr::mutate(gt_AR = if_else(gt_DP!=0 & !is.na(gt_DP), gt_alt.AD/gt_DP, as.numeric(NA))) %>%
-           dplyr::mutate(TYPE = if_else(str_length(REF)>1 || str_length(ALT)>1, "INDEL", "SNV")) %>%
-           dplyr::filter(!Indiv %in% c("SOJ-2-6-2-2-bis", "SOJ-5-3-3-1-bis", "SOJ-3-5-3-1"))
+```{r}
 
-# la taille du tableau
-#dim(vcf.tbl)
-#head(vcf.tbl)
+vcfR <- vcfR::read.vcfR(vcf.file, verbose = FALSE)
 
-samples <- unique(vcf.tbl$Indiv)
+vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE)
 
-samples
+vcfR.tb <- full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS")) %>% 
+  dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))
+
+# split multiallelic variants + add columns  with var type information
+variants.gatk <- vcfR_split_multiAlt(vcfR.tb) %>% 
+  dplyr::mutate(var_length = str_length(ALT) - str_length(REF),
+                var_type   = dplyr::if_else(var_length == 0, "SNV", 
+                             dplyr::if_else(var_length  < 0, "DEL", "INS")),
+                ID = paste(CHROM, POS, var_type, var_length, sep = "_"),
+                tools = "GATK")
+
+#!!!!!! CA    AA   
 ```
 
-# contrôle qualité et nettoyage des données
+## Quality control and filtering
 
-## Les métriques de qualité
+> Before Filtering
 
-```{r, score quality, out.width="50%", fig.show="hold", message=FALSE, warning=FALSE, echo=FALSE}
+```{r score quality, out.width="50%", fig.show="hold", message=FALSE, warning=FALSE, echo=FALSE}
 
 # Score de qualité
-p <- ggplot(data=vcf.tbl) + geom_density(aes(QUAL)) + 
-          xlab("score de qualité") + ylab("densité") + xlim(c(0,200))
+p <- variants.gatk %>% select(ID, QUAL) %>% unique() %>% ggplot() + geom_density(aes(QUAL)) + xlim(0,200) +
+  xlab("quality score") + geom_vline(xintercept = qual.min, color = "red") + ggtitle("Variant quality score distribution")
 print(p)
 
+vcf.tbl.g   <- dplyr::filter(variants.gatk,   gt_AR != 0, !is.na(gt_AR))
+
 # distribution de la couverture des positions variantes
-p <- ggplot(data=vcf.tbl) + geom_boxplot(aes(y=gt_DP, x=Indiv, color=Indiv), outlier.shape = NA) + 
-          ylab("DP of variants") + ylim(0, 100) +
-          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(x=gt_DP, color=Indiv)) + xlim(0,50) +
+  ylab("DP of variants") + geom_vline(xintercept = DP.min, color = "red") + ggtitle("Variant depth of coverage distribution")
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
+
 print(p)
@@
 # Distribution de l'allèle ratio
-p <- ggplot(data=vcf.tbl) + geom_density(aes(gt_AR, color=Indiv)) + 
-          xlab("Ratio de l'allèle alternatif") + ylab("densité")
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(gt_AR, color=Indiv)) + 
+  xlab("Alternative allele ratio") + xlim(0,1) + geom_vline(xintercept = AR.min, color = "red")+ ggtitle("Variant allele ratio distribution")
 
-if (length(samples) > 15){
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
   p <- p + theme(legend.position="none")
@@
 # compter le nombre de positions non couvertes par échantillon.
-NA_count_samples <- dplyr::select(vcf.tbl, ID, Indiv, gt_DP) %>% 
-                    dplyr::group_by(Indiv) %>% dplyr::count(is.na(gt_DP)) %>%
-                    dplyr::filter(`is.na(gt_DP)` == TRUE)
-
-p <- ggplot(data=NA_count_samples) +  geom_bar(aes(y=n, x=Indiv), stat = "identity") + 
+p <- ggplot(data = vcf.tbl.g) +  geom_bar(aes(x=Indiv)) + 
   theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("") + 
-  ylab("Positions manquées (non couvertes)")
+  ylab("Non covered positions")+ ggtitle("Non covered variant positions")
 print(p)
@@
 
-## Filtres qualité
+```{r var depth, out.width="50%", fig.show="hold"}
+NA_ID <- filter(variants.gatk, nb_NA > 0)$ID
+
+variants.gatk %>% filter(ID %in% NA_ID) %>% 
+  ggplot() +  geom_bar(aes(x=as.factor(nb_NA))) + 
+  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("") + 
+  ylab("Non covered positions") + ggtitle("Non covered variant positions")
+
+```
+
+> filtering threshold
+
 * score de qualité : > `r qual.min`
@@
 * nombre de reads qui portent l'allèle alternatif : > 1
+* min AR : `r AR.min`
 * retirer les positions avec au moins un échantillon non couvert.
 
-```{r filtre qual, message=FALSE, warning=FALSE, echo=FALSE}
+```{r qc filtre}
 
- vcf.tbl.q <- dplyr::filter(vcf.tbl, QUAL >= qual.min) %>%
-              dplyr::mutate(gt_alt.AD = if_else(gt_alt.AD < 2, 0, as.numeric(gt_alt.AD)),
-                            gt_DP     = if_else(gt_DP < DP.min, as.numeric(NA), as.numeric(gt_DP))) %>%
-              dplyr::mutate(gt_AR = if_else(gt_DP!=0 & !is.na(gt_DP), gt_alt.AD/gt_DP, as.numeric(NA)))
+# filter variants with low quality/coverage + add amplicon positions
+variants.gatk <- filter(variants.gatk, nb_NA == 0)
+variants.gatk.flt <- qc_filtering(variants.gatk, DP.min, AR.min, qual.min)
 
-#dim(vcf.tbl.q)
-
-# vcf.tbl.count <- vcf.tbl %>% select(ID, gt_AR, Indiv) %>%  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
-#   dplyr::mutate(NA_count = rowSums(is.na(dplyr::select(., -ID)))) %>% 
-#   dplyr::mutate(count_var = rowSums(0 != (dplyr::select(., -ID, -NA_count)))) %>% 
-#   select(ID, NA_count, count_var)
-
-vcf.tbl.q.count <- vcf.tbl.q %>% select(ID, CHROM, gt_AR, Indiv) %>%  
-  reshape2::dcast(CHROM+ID~Indiv, value.var = "gt_AR") %>%
-  dplyr::mutate(NA_count.q = rowSums(is.na(dplyr::select(., -ID)))) %>% 
-  dplyr::mutate(count_var.q = rowSums(0 != (dplyr::select(., -ID, -NA_count.q)))) %>% 
-  dplyr::filter(NA_count.q == 0 , !is.na(count_var.q))
-
-vcf.tbl.q <- vcf.tbl.q %>% filter(ID %in% vcf.tbl.q.count$ID)
-
-#full_join(vcf.tbl.q.count , vcf.tbl.count , by="ID") %>% filter(NA_count != NA_count.q)
+# dplyr::select(CHROM, POS, ID, REF, ALT, QUAL, EFF, Indiv, gt_GT, gt_AD, gt_DP, gt_AR, 
+#               var_length, var_type,  amplicon, locus, strand, freq, tools)
 
@@
 
-# Statistiques descriptives
+## Descriptive statistics
 
-## Nombre des variants par échantillon.
+### Number of small variants
 
-```{r, out.width="90%", message=FALSE, warning=FALSE, echo=FALSE}
-vcf.tbl.q %>% dplyr::filter(gt_AR > 0) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
-  unique() %>% dplyr::group_by(Indiv) %>% dplyr::count(name="nbr_var") %>% 
-  ggplot() + geom_bar(aes(x=Indiv, y=nbr_var, fill=Indiv), stat = "identity") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") + 
-  xlab("") + ylab("# variants")
+```{r nb_SNV, out.width="90%"}
+
+vcf.tbl.q.g <- filter(variants.gatk.flt, gt_AR != 0, !is.na(gt_AR))
+vcf.tbl.q.g %>% dplyr::filter(freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv, var_type) %>% unique() %>% 
+  ggplot() + geom_bar(aes(x=Indiv, fill=var_type)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
+  xlab("") + ylab("nb of variants")
 ```
 
-## Frequences alléliques dans les échantillons.
+### Variant frequency
 
-```{r, out.width="60%", message=FALSE, warning=FALSE, echo=FALSE}
-# retirer les lignes variant-echantillon avec un allèle ref. 
-vcf.tbl.q.var <- dplyr::filter(vcf.tbl.q, gt_AR > 0)
+```{r SNV_freq, out.width="90%"}
 
-# Découper la la ligne et dire ce que chacune fait?
-ID.uniq <- vcf.tbl.q.var %>% dplyr::select(ID, Indiv) %>% dplyr::group_by(ID) %>% count()
-
-```
-
-```{r, out.width="90%", message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
-df <- dplyr::full_join(ID.uniq, vcf.tbl.q.var, by="ID") %>% 
-      dplyr::select(Indiv, n) %>% dplyr::group_by(n, Indiv) %>% dplyr::count()
-
-p <- ggplot(data=df, aes(x=as.factor(n), y=nn)) + geom_boxplot(outlier.shape = NA) +
+p <- vcf.tbl.q.g %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
+  dplyr::filter(freq != length(unique(sample_sheet$sample))) %>% dplyr::count(name = "nb_var") %>% 
+  ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
   geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")
 
-if (length(samples) > 15){
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
 
@@
 print(p)
+
+vcf.tbl.g %>% filter(POS == 1548127)
+
+# warning -> samples with the same variants or pas de variant
 ```
 
-# recherche de pics
+### Clustering
 
-* points gris : variations ref ou fréquentes (communes à 5 échantillons et plus).
-* points verts : variations non ref et non fréqunetes.
-* points rouges : variations non ref et non fréquentes avec un ratio égales ou sup à `r hom`.
+```{r SNV_clustering, out.width="90%"}
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
-vcf.tbl.q.var <- dplyr::filter(vcf.tbl.q, gt_AR > 0, !is.na(gt_AR))
+df <- vcf.tbl.q.g %>% dplyr::filter( freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv , value.var = "gt_AR")
 
-ID.ref  <- dplyr::filter(vcf.tbl.q.var, Indiv == ref) 
-#vcf.tbl.non.ref <- dplyr::filter(vcf.tbl.q.var, ! ID %in% ID.ref$ID)
+# calcul la distance entre les lignes de la matrice
+# vcf.tidy.bin[,-1] : sans la 1ere colonne qui contient les IDs
+dist  <- dist(t(df[,-1]), method = "euclidean")
 
-#dim(vcf.tbl.non.ref)
+# clustering hiérarchique
+hc    <- hclust(dist)
+
+# arranger la structure du dendrogramme
+dend  <- as.dendrogram(hc)
+
+# dessiner le dendrogramme
+plot(dend, main = "Sample clustering using all variants")
 ```
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
+### 3 type de substitution for AG > TA ...
 
-# test sans dplyr::select -> error pour illuster pourquoi il est important de rajouter le nom du package avnat la fonction
-#ID.uniq <- select(vcf.tbl.non.ref, ID, Indiv) %>% group_by(ID) %>% count() %>% filter(n < 2) 
-ID.uniq <- dplyr::select(vcf.tbl.q.var, ID, Indiv) %>% group_by(ID) %>% count() %>% filter(n < 2)
+```{r}
 
-vcf.tbl.uniq <- dplyr::filter(vcf.tbl.q.var, ID %in% ID.uniq$ID)
-vcf.tbl.nonref  <- dplyr::filter(vcf.tbl.q.var, !ID %in% ID.ref$ID, ID %in% filter(vcf.tbl.q.count, count_var.q < 5)$ID)
-#
-#dim(vcf.tbl.uniq)
+p <- vcf.tbl.q.g %>% dplyr::filter(var_type == "SNV", str_length(ALT) == 1) %>% 
+  dplyr::select(ALT, REF, Indiv) %>% group_by(ALT, REF, Indiv) %>% 
+  dplyr::count(name = "nb_sub") %>% 
+  ggplot(aes(x=paste(REF, ALT, sep=">"), y=nb_sub)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2))
+
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
+
+  p <- p + theme(legend.position="none")
+}
+print(p)
 
@@
 
-```{r plot all indiv, warning=FALSE, message=FALSE, fig.height=2, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
+### 4 annotations
+```{r}
 
-mutations.list <- list()
+colname.ANN <- c("Allele", "effect", "impact", "GeneName", 
+                 "GeneID", "FeatureType", "FeatureID", "TranscriptBiotype", 
+                 "Rank", "HGVS.c", "HGVS.p", "cDNA_len:", 
+                 "CDS_len", "Protein_len", "DistanceToFeature", "Warnings")
 
-# une boucle sur l'ensemble des échantillons
-for(ind in unique(vcf.tbl$Indiv)){
+exlude_effect <- c("upstream_gene_variant", "downstream_gene_variant", "intergenic_region", "intragenic_variant", "intragenic_variant", 
+                   "3_prime_UTR_variant", "", "", "")
 
-  print(ind)
+ID.ANN <- vcf.tbl.q.g %>% dplyr::select(ID, ANN) %>% unique()
 
-#  vcf.tbl.ind         <- dplyr::filter(vcf.tbl.q,       Indiv==ind)
-  vcf.tbl.q.var.ind   <- dplyr::filter(vcf.tbl.q.var,   Indiv==ind)
-  #vcf.tbl.uniq.ind    <- dplyr::filter(vcf.tbl.uniq,    Indiv==ind)
-  vcf.tbl.uniq.ind  <- dplyr::filter(vcf.tbl.nonref,     Indiv==ind)
+ANN.split <- lapply(1:dim(ID.ANN)[1], FUN = function(x){
 
-  hom.chr <- dplyr::filter(vcf.tbl.uniq.ind, gt_AR > hom) %>% dplyr::select(CHROM) %>% unique()
-  var.to.smooth <- dplyr::filter(vcf.tbl.uniq.ind, CHROM %in% hom.chr$CHROM, gt_AR >= AR.min)
+  tmp <- ID.ANN$ANN[x] %>% str_split(",") %>% unlist()
+  data.frame(ID = rep(ID.ANN$ID[x], length(tmp)), ANN = tmp)
 
-  # stocker le res dans une list
-  mutations.list[[ind]] <- var.to.smooth %>% dplyr::filter(gt_AR > hom)
+}) %>% purrr::reduce(rbind) %>% 
+  tidyr::separate(col = ANN, into = colname.ANN, sep = "[|]") %>% 
+  dplyr::filter(!effect %in% exlude_effect) %>% 
+  dplyr::select(ID, effect) %>% unique()
 
-  p <- ggplot() + ylim(0,1) +
-      geom_point( data=vcf.tbl.q.var.ind,    aes(x=POS, y=gt_AR), size=0.2, color="grey80") + 
-      geom_point( data=vcf.tbl.uniq.ind,     aes(x=POS, y=gt_AR), size=0.3, color="green") + 
-      geom_point( data=mutations.list[[ind]],aes(x=POS, y=gt_AR), size=0.5, color="red") + 
-      geom_smooth(data=var.to.smooth,        aes(x=POS, y=gt_AR), size=0.5) +
-      facet_grid( ~ CHROM, scale="free",  space = "free") +
-      theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0),
-            axis.text.y = element_text(size = rel(0.8), hjust = 0)) +
-      scale_x_continuous(breaks = seq(0, max(vcf.tbl$POS), 1e+07))
-  print(p)
+effect.split <- lapply(1:dim(ANN.split)[1], function(x){
 
-}
+  tmp <- ANN.split$effect[x] %>% stringr::str_split("&") %>% unlist()
+  data.frame(ID = rep(ANN.split$ID[x], length(tmp)), effect = tmp)
+}) %>% purrr::reduce(rbind)
 
-# concatener les tibbles
-mutations <- purrr::reduce(mutations.list, rbind) %>% select(CHROM, ID, QUAL, gt_DP, gt_ref.AD, gt_alt.AD, gt_AR, Indiv, ANN) %>% separate(ANN, c("tmp1","eff", "tmp2", "gene"), "\\|") %>% select(-paste0("tmp",1:2))
-
-mutations %>% select(Indiv, ID, eff) %>% filter(!eff %in% c("downstream_gene_variant", "intergenic_region", "upstream_gene_variant")) %>% unique() %>% group_by(Indiv, eff) %>% count(name = "count")
-
-write.table(mutations, file="SOJ_filt2_mutation_Ho.xls", quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t")
+effect.split %>% dplyr::group_by(effect) %>% dplyr::count()
 

### 2023-06-20 script/quality_control.Rmd
update report

@@
-## Trimming parameters
+# Quality control
+## Read quality
+
+<!-- <details> -->
+<!--     <summary> multiqc report </summary> -->
+
+<!-- ```{r multiqc} -->
+
+<!-- htmltools::includeHTML(paste0(workspace,"/01_sequence_qc/SOJ_multiqc_trim.html")) -->
+
+<!-- ``` -->
+<!-- </details> -->
+
+### Read cleaning :
+
+Sequences were trimmed with Trimmomatic tools...
 
 ```{r trimmomatic}
-trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
-dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
-DT::datatable(dat)
+trimmo <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
+dat    <- read.table(text = gsub(" ","\t",readLines(trimmo)),  header=TRUE)
+DT::datatable(dat,  rownames = FALSE, caption = "Read cleaning results",
+              options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
 ```
 
-## HTML
+## Alignement
 
-## Mapping parameters
+Sequences were aligned using BWA MEM (v?) ...
 
-### Flagstat results
+### Mapping rate
 
-```{r }
-
+```{r mapping rate}
 file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
@@
     flagstat <- read.table(file, sep="\t", header=FALSE) %>%
-      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
-                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
-                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
-                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
-    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")
+      dplyr::mutate(mapped=V3-V4, mapped.ratio=round(mapped/(V2)*100,2)) %>% dplyr::select(-V3, -V4)
+    names(flagstat) <- c("sample", "raw", "mapped", "mapped.ratio")
 
@@
 
-DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
+DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe',
+              rownames = FALSE, caption = "???")
+
+```
+
+```{r}
 
 #count
-nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
+nbr.reads.melt <- flagstat %>% dplyr::select(sample, raw, mapped, mapped.ratio) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
 
-p <- ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=variable)) + geom_boxplot()  + 
-  xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
+# p <- ggplot(data = dplyr::filter(nbr.reads.melt, variable == "mapped.ratio"), aes(y=nbr.reads, x=variable)) + geom_boxplot()  +
+#   xlab("") + ylab("% reads") + ggtitle("...")
+# 
+# print(p)
+# 
+# p <- ggplot(data = dplyr::filter(nbr.reads.melt, variable %in% c("raw", "mapped")), aes(y=nbr.reads, x=variable)) + geom_boxplot()  +
+#   xlab("") + ylab("number of reads") + ggtitle("...")+theme_bw()
+# 
+# print(p)
 
-print(p)
-
-#ratio
-ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
-names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
-
-ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
-
-summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% 
-  summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
-
-p <- ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=variable)) + 
-  xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + 
-  ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()
+p <- ggplot(data = dplyr::filter(nbr.reads.melt, variable %in% c("mapped.ratio", "mapped")), aes(x=nbr.reads, y=variable)) + geom_boxplot()  +
+  xlab("") + ylab("") + ggtitle("Mapped reads per samples") + facet_wrap(variable~., scales = "free", ncol = 1) +
+  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
 
@@
 
-### Coverage analysis
-```{r coverage}
+### Coverage
+
+```{r depth}
+
 cov.list <- lapply(sample_sheet$sample, function(x){
@@
 
-cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
-colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
-#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
-#cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
-#cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
+cov <- purrr::reduce(cov.list, rbind)
+colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample")
+cov$coverage <- round(cov$coverage, 2)
 
-p <-ggplot(data = cov) + geom_boxplot(aes(y=meandepth, x=rname)) + 
-  xlab("") + ggtitle("Plot coverage")  + 
-  theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()
+p <-ggplot(data = cov) + geom_boxplot(aes(x=meandepth, y=rname)) + xlab("depth (X)") + ylab("") + ggtitle("Depth of coverage mean per samples") +
+  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) + 
+  facet_wrap(rname~., scales = "free")
+
+print(p)
+
+p <-ggplot(data = cov) + geom_boxplot(aes(x=coverage, y=rname)) +  ylab("") + xlab("Coverage (%)") + ggtitle("Propotion of covered bases per sample") +
+  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) + scale_x_continuous(labels=function(x){sprintf("%.2f", x)}) + facet_wrap(rname~., scales = "free")
 

### 2023-06-20 script/report.Rmd
update report

@@
     html_document:
-        highlight: tango
-        number_sections: no
-        theme: default
-        toc: yes
-        toc_depth: 3
-        toc_float:
-            collapsed: no
-            smooth_scroll: yes
+        number_sections: yes
 params:
@@
   AR.min: AR.min
+  qual.min: qual.min
   result_dir: result_dir
+  sample_sheet: sample_sheet
   rmd: "report.Rmd"
@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE)
+```
+
+```{r}
 # title: "`r params$title`"
-workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
-#workspace<-params$result_dir
+# workspace<-params$result_dir
+# sample_sheet.file <- params$sample_sheet
+# DP.min <- params$DP.min
+# AR.min <- params$AR.min
+# qual.min <- params$qual.min
 
-sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+projectName <- "PE-WGS"
 
-knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=TRUE, message=FALSE)
+workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
+#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
+
+sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
+#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+
+DP.min <- 4
+AR.min <- 0.8
+qual.min <- 30
+
+# results
+trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
+vcf.file <- paste0(workspace,"/03_snv_indels_calling/", projectName , "_snv_indel.vcf")
+sv.file  <- paste0(workspace,"/04_sv_calling/", projectName , "_sv.vcf")
+
+if(!file.exists(sample_sheet.file)) stop("Error: ")
+if(!file.exists(trim.file)) stop("Error: ")
+if(!file.exists(vcf.file)) stop("Error: ")
+
+eval.SV <- file.exists(sv.file)
+
 ```
@@
 library(dplyr)
+library(tidyr)
 library(yaml)
+library(stringr)
+library(ComplexHeatmap)
+library(gplots)
+source("functions.R")
 
@@
 
+# Samples
 ```{r}
@@
 
+unique(sample_sheet$sample)
+```
+List of samples : `r unique(sample_sheet$sample)`
+
+```{r}
+
+dplyr::select(sample_sheet, sample, condition) %>% unique() %>% dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
+  knitr::kable(align = "lccrr")
+ # DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
+
 ```
 
-```{r quality_control.Rmd}
+<!-- # Quality control -->
+
+```{r quality_control}
 
@@
 
-```{r SNV_InDels_calling.Rmd, eval=FALSE}
+<!-- # SNV and Small InDels Detection -->
 
-vcf.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/03_snv_indels_calling/SOJ_snv_indel.vcf"
+```{r SNV_InDels_calling}
 
@@
 
-```{r SV_calling.Rmd, eval=FALSE}
+<!-- # SV Detection -->
+
+```{r SV_calling, eval=eval.SV}
 
@@
 
-```{r tDNA_scan.Rmd, eval=FALSE}
+```{r tDNA_scan, eval=FALSE, include=FALSE}
 

### 2023-06-20 script/functions.R

@@ -1,0 +1,104 @@
+vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){
+  
+  vcfR.tid.tmp <- 
+    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), 
+                                                    function(x){return(length(x))}))) %>%
+    dplyr::mutate(nb.allele=nb.allele-1)
+  
+  vcfR.tid.spt <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){
+    
+    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
+      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = ALT,   into = paste0("ALT"  ,1:x), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = AC,    into = paste0("AC"   ,1:x), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = AF,    into = paste0("AF"   ,1:x), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = MLEAF, into = paste0("MLEAF",1:x), sep = ",", remove = FALSE) %>%
+      tidyr::separate(col = MLEAC, into = paste0("MLEAC",1:x), sep = ",", remove = FALSE)
+    
+    tmp.list <- lapply(1:x, FUN = function(i){ 
+      
+      tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
+                                 gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
+                                 ID_old    = ID,
+                                 gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
+                                 ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
+                                 AC_old    = AC,    AC    = get(paste0("AC", i)),
+                                 AF_old    = AF,    AF    = get(paste0("AF", i)),
[…]

### 2023-07-06 script/report.Rmd
update report

--- 2023-04-11 script/report.Rmd
@@
         highlight: tango
-        number_sections: no
+        number_sections: yes
         theme: default
@@
         toc_float:
-            collapsed: no
+            collapsed: yes
             smooth_scroll: yes
@@
 
-```{r setup, include=FALSE}
+```{r}
 # title: "`r params$title`"
-workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
-#workspace<-params$result_dir
+workspace <-params$result_dir
+sample_sheet.file <- params$sample_sheet
+# DP.min <- params$DP.min
+# AR.min <- params$AR.min
+# qual.min <- params$qual.min
 
-sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
+projectName <- "PE-WGS"
 
-knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=TRUE, message=FALSE)
+#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
+#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
+
+#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
+#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+
+DP.min <- 4
+AR.min <- 0.8
+qual.min <- 30
+nbr.alt.min <- 3
 ```
 
-```{r }
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
+```
+
+```{r}
+#libraries
 library(DT)
@@
 library(dplyr)
+library(tidyr)
 library(yaml)
+library(stringr)
+library(ComplexHeatmap)
+library(dendextend) 
+library(GenomicRanges)
+library(gplots)
+source("functions.R")
+```
 
+# Objective
+
+# Input data
+## Genome version
+```{r, eval=FALSE}
+print(basename(Genome.file))
+```
+
+## Samples
+```{r}
+# error message if sample sheet not exist
+if(!file.exists(sample_sheet.file)) stop("Error: ")
+
+# read sample sheet
+sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
+```
+
+```{r}
+sampleNames <- unique(sample_sheet$sample)
+```
+
+```
+List of samples : `r sampleNames`
+```
+
+```{r list samples}
+
+dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
+  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe', caption = "List of conditions")
+
+```
+
+# MsGenoVa workflow
+
+![alt text here](workflow.png)
+
+# Sequence trimming step
+
+Sequences were trimmed with Trimmomatic tools...
+
+```{r trimming stat}
+trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
+if(!file.exists(trim.file)) stop("Error: ")
+
+read.table(text = gsub(" ","\t",readLines(trim.file)),  header=TRUE) %>%
+  DT::datatable(rownames = FALSE, caption = "Read cleaning results",
+                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe')
+
+```
+
+# Mapping step
+
+Sequences were aligned using BWA MEM (v?) ...
+
+```{r mapping step}
+map.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(!file.exists(map.file)) stop("Error: ")
+```
+
+## Mapping rate
+
+```{r mapping rate}
+flagstat.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(file.exists(flagstat.file)){
+    flagstat <- read.table(flagstat.file, sep="\t", header=FALSE) %>%
+      dplyr::mutate(mapped.reads=V3-V4, mapped.read.ratio=round(mapped.reads/(V2)*100,2)) %>% dplyr::select(-V3, -V4)
+    names(flagstat) <- c("sample", "raw.reads", "mapped.reads", "mapped.read.ratio")
+
+  }
+
+#count
+nbr.reads.melt <- flagstat %>% dplyr::mutate(xlabel=paste0(sample, "(", mapped.read.ratio, "%)"), unmapped.reads=raw.reads-mapped.reads) %>% 
+  dplyr::select(mapped.reads, unmapped.reads, xlabel) %>%
+  reshape2::melt(id = "xlabel", value.name = "nbr.reads")
+
+```
+
+```{r mapping rate plot}
+p <- ggplot(data = nbr.reads.melt) + geom_col(aes(y=xlabel, x=nbr.reads, fill=variable)) + 
+  labs(x="nb of reads", y="samples (mapping rate %)", fill="")
+
+print(p)
+```
+
+```{r mapping rate summary}
+DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+              class = 'cell-border stripe', rownames = FALSE, caption = "")
+```
+
+## Mapping coverage
+
+```{r mapping coverage}
+cov.list <- lapply(sample_sheet$sample, function(x){
+  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
+  if(file.exists(file)){
+
+    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
+  }
+})
+
+cov <- purrr::reduce(cov.list, rbind)
+colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage(%)",	"meandepth(X)",	"meanbaseq",	"meanmapq", "sample")
+cov$coverage <- round(cov$coverage, 2)
+
+cov <- cov %>% mutate(xlabel=paste0(rname, "(", endpos-startpos, ")")) %>% 
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "scaffold")) %>%
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "chloroplast")) %>%
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "mitochondria"))
+```
+
+```{r mapping coverage plot}
+p <- select(cov, rname, `meandepth(X)`, `coverage(%)`, sample, xlabel) %>% reshape2::melt(id = c("rname", "sample", "xlabel")) %>% 
+  ggplot() + geom_boxplot(aes(x=value, y=xlabel)) + 
+  xlab("") + ylab("chromosome names (size bp)") + 
+  facet_grid(~variable, scales = "free")
+
+print(p)
+```
+```{r mapping coverage summary}
+cov %>% dplyr::select(sample, rname, `meandepth(X)`, `coverage(%)`) %>% 
+  DT::datatable(rownames = FALSE, caption = "",
+                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe')
+```
+
+# Small variant calling step (GATK)
+```{r}
+vcf.file <- paste0(workspace,"/03_snv_indels_calling/", projectName , "_snv_indel.vcf")
+if(!file.exists(vcf.file)) stop("Error: ")
 ```
@@
 
-sample_sheet <- read.table(file = sample_sheet.file, header = TRUE, sep = "\t")
+vcfR <- vcfR::read.vcfR(vcf.file, verbose = FALSE)
+
+vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE, info_fields = c("AC","MQ", "ANN"), 
+                                                     format_fields = c("Indiv", "AD", "DP"))
+
+vcfR.tb <- dplyr::full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS"))
+rm(vcfR.tb.list)
+
+# split multiallelic variants + add columns  with var type information
+variants.gatk <- vcfR_split_multiAlt(vcfR.tb, nbrAlt = 2)
+
+variants.gatk <- variants.gatk %>%
+  tidyr::separate(gt_AD, c("gt_ref_AD", "gt_alt_AD"), sep = ",") %>%
+  dplyr::mutate(gt_AR = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP),
+                var_length = str_length(ALT) - str_length(REF) + 1,
+                var_type   = dplyr::if_else(var_length == 1, "SNV", 
+                             dplyr::if_else(var_length  < 1, "DEL", "INS")),
+                ID = paste(CHROM, POS, REF, ALT, sep = "_"), tools = "GATK")
+
+variants.gatk <- variant.freq(variants.gatk)
 
@@
 
-```{r quality_control.Rmd}
+## Variant calling quality control
 
-  out <- NULL
+### Score quality
 
-  out <- c(out, knitr::knit_child("quality_control.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
+```{r out.width="60%", fig.show="hold"}
+# Score de qualité
+p <- variants.gatk %>% select(ID, QUAL) %>% unique() %>% ggplot() + geom_density(aes(QUAL)) + xlim(0,200) +
+  xlab("quality score") + geom_vline(xintercept = qual.min, color = "red") + ggtitle("Variant quality score distribution")
+print(p)
 
@@
 
-```{r SNV_InDels_calling.Rmd, eval=FALSE}
+### Variant coverage
 
-vcf.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/03_snv_indels_calling/SOJ_snv_indel.vcf"
+```{r out.width="50%", fig.show="hold"}
+vcf.tbl.g   <- dplyr::filter(variants.gatk,   gt_AR != 0, !is.na(gt_AR))
 
-  out <- NULL
+# distribution de la couverture des positions variantes
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(x=gt_DP, color=Indiv)) + xlim(0,50) +
+  ylab("DP of variants") + geom_vline(xintercept = DP.min, color = "red") + ggtitle("Variant depth of coverage distribution")
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
 
-  out <- c(out, knitr::knit_child("SNV_InDels_calling.Rmd"), quiet=TRUE)
+print(p)
 
-knitr::asis_output(out)
+# Distribution de l'allèle ratio
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(gt_AR, color=Indiv)) + 
+  xlab("Alternative allele ratio") + xlim(0,1) + geom_vline(xintercept = AR.min, color = "red")+ ggtitle("Variant allele ratio distribution")
+
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
+
+print(p)
 
@@
 
-```{r SV_calling.Rmd, eval=FALSE}
+```{r out.width="60%", fig.show="hold"}
+NA_ID <- filter(variants.gatk, nb_NA > 0)$ID
 
-  out <- NULL
-
-  out <- c(out, knitr::knit_child("SV_calling.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
+variants.gatk %>% filter(ID %in% NA_ID) %>% 
+  ggplot() +  geom_bar(aes(x=as.factor(nb_NA))) + 
+  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("nb of samples") + 
+  ylab("Non covered positions") + ggtitle("Non covered variant positions")
 
@@
 
-```{r tDNA_scan.Rmd, eval=FALSE}
+### Low quality variant filtering
 
-  out <- NULL
+* score de qualité : > `r qual.min`
+* nombre de reads qui couvrent la position variant : > `r DP.min`
+* nombre de reads qui portent l'allèle alternatif : > 1
+* min AR : `r AR.min`
+* retirer les positions avec au moins un échantillon non couvert.
 
-  out <- c(out, knitr::knit_child("tDNA_scan.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
+```{r}
+# remove 
+variants.gatk.flt <- variants.gatk %>% 
+  dplyr::mutate(gt_alt_AD = dplyr::if_else(as.numeric(gt_alt_AD) < 2, 0, as.numeric(gt_alt_AD)),
+                gt_AR     = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP)) %>% 
+  dplyr::filter(!is.na(gt_DP), gt_DP >= DP.min, QUAL >= qual.min, nb_NA == 0, freq_ho == freq)
 
@@
 
-```{r tools, eval=FALSE}
+## Descriptive stat of variants
 
-  out <- NULL
+### variant count per sample
 
-  out <- c(out, knitr::knit_child("tools.Rmd"), quiet=TRUE)
+```{r }
 
-knitr::asis_output(out)
+vcf.tbl.q.g <- filter(variants.gatk.flt, gt_AR != 0, !is.na(gt_AR))
+
+vcf.tbl.q.g %>% 
+  dplyr::mutate(tag=if_else(freq == length(unique(sample_sheet$sample)), "common", "noCommon")) %>% 
+  dplyr::select(ID, Indiv, var_type, tag) %>%
+  dplyr::mutate(Indiv=if_else(tag == "common", "common", Indiv)) %>% unique() %>%
+  ggplot() + geom_bar(aes(x=Indiv, fill=var_type)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
+  xlab("") + ylab("nb of variants")
 
 ```
+* common : les variants communs à tous les échantillons.
+
+### Variant frequency in samples
+
+```{r }
+
+p <- vcf.tbl.q.g %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
+  dplyr::count(name = "nb_var")  %>% filter(freq != length(unique(sample_sheet$sample))) %>% 
+  ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")
+
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
+
+  p <- p + theme(legend.position="none")
+}
+print(p)
+
+# warning -> samples with the same variants or pas de variant
+```
+
+### Substitution
+
+```{r}
+
+p <- vcf.tbl.q.g %>% dplyr::filter(var_type == "SNV", str_length(ALT) == 1) %>% 
+  dplyr::select(ALT, REF, Indiv) %>% group_by(ALT, REF, Indiv) %>% 
+  dplyr::count(name = "nb_sub") %>% 
+  ggplot(aes(x=paste(REF, ALT, sep=">"), y=nb_sub)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + 
+  facet_grid(~paste0("REF : ",REF), scales = "free") + labs(x="", y="nb of variants", color="samples")
+
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
+
+  p <- p + theme(legend.position="none")
+}
+print(p)
+
+```
+
+### Clustering per condition
+
+```{r}
+vcf.tbl.q.g.condition <- vcf.tbl.q.g %>% 
+  dplyr::full_join(dplyr::select(sample_sheet, sample, condition), by = c("Indiv" = "sample")) %>% 
+  tidyr::separate(ANN, c("tmp1", "effect", "tmp2", "gene"), sep = "[|]", extra = "drop") %>% dplyr::select(-tmp1, -tmp2)
+
+tb <- vcf.tbl.q.g.condition %>%  
+  dplyr::select(ID, condition) %>% 
+  dplyr::group_by(ID, condition) %>% unique() %>% 
+  reshape2::dcast(ID~condition, value.var = "condition") %>%
+  dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)})
+
+venn::venn(tb[,-1])
+```
+
+### Clustering per sample
+
+```{r }
+
+df <- vcf.tbl.q.g %>% dplyr::filter( freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv , value.var = "gt_AR")
+
+# calcul la distance entre les lignes de la matrice
+# vcf.tidy.bin[,-1] : sans la 1ere colonne qui contient les IDs
+dist  <- dist(t(df[,-1]), method = "euclidean")
+
+# clustering hiérarchique
+hc    <- hclust(dist)
+
+# arranger la structure du dendrogramme
+dend  <- as.dendrogram(hc)
+
+col.rainbow <- colorspace::rainbow_hcl(length(unique(sample_sheet$condition)))
+
+col <- col.rainbow[as.factor(sample_sheet$condition)]
+names(col) <- sample_sheet$sample
+
+labels_colors(dend) <- col[order.dendrogram(dend)]
+# dessiner le dendrogramme
+plot(dend, main = "Sample clustering using all variants")
+```
+
+```{r, eval=FALSE}
+## annotations
+
+colname.ANN <- c("Allele", "effect", "impact", "GeneName", 
+                 "GeneID", "FeatureType", "FeatureID", "TranscriptBiotype", 
+                 "Rank", "HGVS.c", "HGVS.p", "cDNA_len:", 
+                 "CDS_len", "Protein_len", "DistanceToFeature", "Warnings")
+
+exlude_effect <- c("upstream_gene_variant", "downstream_gene_variant", "intergenic_region", "intragenic_variant", "intragenic_variant", 
+                   "3_prime_UTR_variant", "", "", "")
+
+ID.ANN <- vcf.tbl.q.g %>% dplyr::select(ID, ANN) %>% unique()
+
+ANN.split <- lapply(1:dim(ID.ANN)[1], FUN = function(x){
+
+  tmp <- ID.ANN$ANN[x] %>% str_split(",") %>% unlist()
+  data.frame(ID = rep(ID.ANN$ID[x], length(tmp)), ANN = tmp)
+
+}) %>% purrr::reduce(rbind) %>% 
+  tidyr::separate(col = ANN, into = colname.ANN, sep = "[|]") %>% 
+  dplyr::filter(!effect %in% exlude_effect) %>% 
+  dplyr::select(ID, effect) %>% unique()
+
+effect.split <- lapply(1:dim(ANN.split)[1], function(x){
+
+  tmp <- ANN.split$effect[x] %>% stringr::str_split("&") %>% unlist()
+  data.frame(ID = rep(ANN.split$ID[x], length(tmp)), effect = tmp)
+}) %>% purrr::reduce(rbind)
+
+effect.split %>% dplyr::group_by(effect) %>% dplyr::count()
+
+```

### 2023-07-06 script/functions.R

@@ -1,40 +1,96 @@
+# vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){
+#   
+#   
+#   vcfR.tid.m <- dplyr::filter(stringr::str_detect(ALT, ",")) %>%
+#     dplyr::mutate(vcfR.tid, nb.allele=stringr::str_count(",")+1)
+#   
+#   vcfR.tid.spt <- lapply(unique(vcfR.tid.m$nb.allele), function(x){
+#     
+#     tmp <- vcfR.tid.m %>% dplyr::filter(nb.allele == x) %>%
+#       tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
+#       tidyr::separate(col = ALT,   into = paste0("ALT"  ,1:x), sep = ",", remove = FALSE) %>%
+#       tidyr::separate(col = AC,    into = paste0("AC"   ,1:x), sep = ",", remove = FALSE) %>%
+#       tidyr::separate(col = AF,    into = paste0("AF"   ,1:x), sep = ",", remove = FALSE) %>%
+#       tidyr::separate(col = MLEAF, into = paste0("MLEAF",1:x), sep = ",", remove = FALSE) %>%
+#       tidyr::separate(col = MLEAC, into = paste0("MLEAC",1:x), sep = ",", remove = FALSE)
+#     
+#     tmp.list <- lapply(1:x, FUN = function(i){ 
+#       
+#       tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
+#                                  gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
+#                                  ID_old    = ID,
+#                                  gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
+#                                  ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
+#                                  AC_old    = AC,    AC    = get(paste0("AC", i)),
+#                                  AF_old    = AF,    AF    = get(paste0("AF", i)),
+#                                  MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
[…]

### 2023-07-06 script/report.Rmd
update report

@@
   sample_sheet: sample_sheet
+  projectName : projectName
   rmd: "report.Rmd"
@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
+```
 
@@
 sample_sheet.file <- params$sample_sheet
-# DP.min <- params$DP.min
-# AR.min <- params$AR.min
-# qual.min <- params$qual.min
-
+DP.min <- params$DP.min
+AR.min <- params$AR.min
+qual.min <- params$qual.min
+projectName <- params$projectName
 #Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
-projectName <- "PE-WGS"
+#projectName <- "PE-WGS"
 
@@
 
-DP.min <- 4
-AR.min <- 0.8
-qual.min <- 30
+# DP.min <- 4
+# AR.min <- 0.8
+# qual.min <- 30
 nbr.alt.min <- 3
-```
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
 ```

### 2023-07-07 script/ms_report.Rmd
debug report

@@
 ---
-title: "MSGenova (Clean version) Rmakrdown Report"
-author: "Gabrièle Adam"
+title: "Report"
+author: " "
 date: "2022-08-16"
@@
   AR.min: AR.min
+  qual.min: qual.min
   result_dir: result_dir
+  sample_sheet: sample_sheet
+  projectName : projectName
   rmd: "report.Rmd"
+  AR: AR
 ---
@@
 ```{r setup, include=FALSE}
-workspace<-params$result_dir
-knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
 ```
 
-```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
-library(tidyverse)
+```{r}
+# title: "`r params$title`"
+workspace <-params$result_dir
+sample_sheet.file <- params$sample_sheet
+DP.min <- params$DP.min
+AR.min <- params$AR.min
+qual.min <- params$qual.min
+projectName <- params$projectName
+#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
+#projectName <- "PE-WGS"
+
+#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
+#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
+
+#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
+#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+
+# DP.min <- 4
+# AR.min <- 0.8
+# qual.min <- 30
+nbr.alt.min <- 3
+```
+
+```{r}
+#libraries
 library(DT)
 library(ggplot2)
-library(plyr)
+library(vcfR)
 library(dplyr)
+library(tidyr)
 library(yaml)
-vector_calling<-params$vector
-
-DP.min<-5
-AR.min<-0.2
-
-sheet<-params$sample
-trimmo<- params$trimmo
-flagstat<- params$flagstat
-haplotype_caller<- params$haplotype_caller
-
-config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
-config_ad<-read_yaml(config_advanced_file)
+library(stringr)
+library(ComplexHeatmap)
+library(dendextend) 
+library(GenomicRanges)
+library(gplots)
+source("functions.R")
 ```
 
-## Trimming parameters
+# Objective
 
-```{r trimmomatic}
-trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
-dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
-DT::datatable(dat)
+# Input data
+## Genome version
+```{r, eval=FALSE}
+print(basename(Genome.file))
 ```
 
-## Mapping parameters
+## Samples
+```{r}
+# error message if sample sheet not exist
+if(!file.exists(sample_sheet.file)) stop("Error: ")
 
-### Flagstat results
+# read sample sheet
+sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
+```
 
-```{r flagstat}
-library(tidyverse)
-file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
-if(file.exists(file)){
-    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
-      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
-                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
-                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
-                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
-    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")
+```{r}
+sampleNames <- unique(sample_sheet$sample)
+```
+
+```
+List of samples : `r sampleNames`
+```
+
+```{r list samples}
+
+dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
+  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe', caption = "List of conditions")
+
+```
+
+# MsGenoVa workflow
+
+![alt text here](workflow.png)
+
+# Sequence trimming step
+
+Sequences were trimmed with Trimmomatic tools...
+
+```{r trimming stat}
+trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
+if(!file.exists(trim.file)) stop("Error: ")
+
+read.table(text = gsub(" ","\t",readLines(trim.file)),  header=TRUE) %>%
+  DT::datatable(rownames = FALSE, caption = "Read cleaning results",
+                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe')
+
+```
+
+# Mapping step
+
+Sequences were aligned using BWA MEM (v?) ...
+
+```{r mapping step}
+map.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(!file.exists(map.file)) stop("Error: ")
+```
+
+## Mapping rate
+
+```{r mapping rate}
+flagstat.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(file.exists(flagstat.file)){
+    flagstat <- read.table(flagstat.file, sep="\t", header=FALSE) %>%
+      dplyr::mutate(mapped.reads=V3-V4, mapped.read.ratio=round(mapped.reads/(V2)*100,2)) %>% dplyr::select(-V3, -V4)
+    names(flagstat) <- c("sample", "raw.reads", "mapped.reads", "mapped.read.ratio")
 
@@
 
-DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
 #count
-nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
-summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
-nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
-ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
-#ratio
-ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
-names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
-ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
-summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
-ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
-ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()
+nbr.reads.melt <- flagstat %>% dplyr::mutate(xlabel=paste0(sample, "(", mapped.read.ratio, "%)"), unmapped.reads=raw.reads-mapped.reads) %>% 
+  dplyr::select(mapped.reads, unmapped.reads, xlabel) %>%
+  reshape2::melt(id = "xlabel", value.name = "nbr.reads")
 
@@
 
-### Coverage analysis
-```{r coverage}
+```{r mapping rate plot}
+p <- ggplot(data = nbr.reads.melt) + geom_col(aes(y=xlabel, x=nbr.reads, fill=variable)) + 
+  labs(x="nb of reads", y="samples (mapping rate %)", fill="")
+
+print(p)
+```
+
+```{r mapping rate summary}
+DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+              class = 'cell-border stripe', rownames = FALSE, caption = "")
+```
+
+## Mapping coverage
+
+```{r mapping coverage}
 cov.list <- lapply(sample_sheet$sample, function(x){
@@
 
-cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
-colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
-#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
-cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
-cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min)) 
-for (i in unique(sample_sheet$sample)){
-cat("####",i,"\n")
-g<-ggplot(filter(cov,sample==i)) + geom_boxplot(aes(y=numreads, x=x.label)) + xlab("") + ylab("#reads") + ggtitle("Plot coverage")+facet_wrap(.
-~ sample,scales = "free")+theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()
-print(g)
-cat('\n\n')
-}
+cov <- purrr::reduce(cov.list, rbind)
+colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage(%)",	"meandepth(X)",	"meanbaseq",	"meanmapq", "sample")
+cov$coverage <- round(cov$coverage, 2)
+
+cov <- cov %>% mutate(xlabel=paste0(rname, "(", endpos-startpos, ")")) %>% 
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "scaffold")) %>%
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "chloroplast")) %>%
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "mitochondria"))
 ```
 
-### Depth analysis
-```{r depth_analysis}
-cov.list <- lapply(sample_sheet$sample, function(x){
+```{r mapping coverage plot}
+p <- select(cov, rname, `meandepth(X)`, `coverage(%)`, sample, xlabel) %>% reshape2::melt(id = c("rname", "sample", "xlabel")) %>% 
+  ggplot() + geom_boxplot(aes(x=value, y=xlabel)) + 
+  xlab("") + ylab("chromosome names (size bp)") + 
+  facet_grid(~variable, scales = "free")
 
-  file <- paste0(workspace,"/02_mapping/depth/", x, ".depth")
-  if(file.exists(file)){
-    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
-  }
-})
+print(p)
+```
+```{r mapping coverage summary}
+cov %>% dplyr::select(sample, rname, `meandepth(X)`, `coverage(%)`) %>% 
+  DT::datatable(rownames = FALSE, caption = "",
+                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe')
+```
 
-depth <- purrr::reduce(cov.list, rbind)
-colnames(depth) <- c("rname", "position", "depth", "sample")
-#depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
-heading<-as.character(unique(sample_sheet$sample))
-for (i in unique(cov$rname)){
- cat("####",i,"\n")
-  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=sample))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free")
-  print(p)
-cat('\n\n')
-}
+# Small variant calling step (GATK)
+```{r}
+vcf.file <- list.files(path = paste0(workspace,"/03_snv_indels_calling"), pattern = "_snv_indel.vcf" , full.names = T)
+if(!file.exists(vcf.file)) stop("Error: ")
+```
+
+```{r}
+
+vcfR <- vcfR::read.vcfR(vcf.file, verbose = FALSE)
+
+vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE, info_fields = c("AC","MQ", "ANN"), 
+                                                     format_fields = c("Indiv", "AD", "DP"))
+
+vcfR.tb <- dplyr::full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS"))
+rm(vcfR.tb.list)
+
+# split multiallelic variants + add columns  with var type information
+variants.gatk <- vcfR_split_multiAlt(vcfR.tb, nbrAlt = 2)
+
+variants.gatk <- variants.gatk %>%
+  tidyr::separate(gt_AD, c("gt_ref_AD", "gt_alt_AD"), sep = ",") %>%
+  dplyr::mutate(gt_AR = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP),
+                var_length = str_length(ALT) - str_length(REF) + 1,
+                var_type   = dplyr::if_else(var_length == 1, "SNV", 
+                             dplyr::if_else(var_length  < 1, "DEL", "INS")),
+                ID = paste(CHROM, POS, REF, ALT, sep = "_"), tools = "GATK")
+
+variants.gatk <- variant.freq(variants.gatk)
 
@@
 
-```{r proportion_of, results='asis',echo=FALSE, eval=FALSE, warning=FALSE, error=FALSE,results=FALSE}
-for (i in unique(bed.tab$target)){
- cat("####",i,"\n")
-tmp <- cov  %>% filter(target== i) %>% group_by(nbr_reads, target, amplicon) %>% count(name = "#well") 
-g1<-ggplot(data = tmp) + geom_bar(aes(x=amplicon, y=`#well`, fill=nbr_reads), stat = "identity") + xlab("") +facet_grid(~target, scales = "free")
-g2<-cov %>% filter(target== i) %>% group_by(well, nbr_reads,target) %>% count(name = "amplicon") %>% filter(nbr_reads == ">15X") %>% group_by(amplicon,target) %>% count(name = "#well") %>% ggplot(aes(x = amplicon, y = `#well`)) + geom_bar(stat = "identity") + xlab("nbr of amplicon covered at >15X")
-print(g1)
-print(g2)
-cat('\n\n')
-}
+## Variant calling quality control
+
+### Score quality
+
+```{r out.width="60%", fig.show="hold"}
+# Score de qualité
+p <- variants.gatk %>% select(ID, QUAL) %>% unique() %>% ggplot() + geom_density(aes(QUAL)) + xlim(0,200) +
+  xlab("quality score") + geom_vline(xintercept = qual.min, color = "red") + ggtitle("Variant quality score distribution")
+print(p)
+
 ```
 
-## Study of SNV and Indels calling
+### Variant coverage
 
-```{r define_r_fct_multiallele, echo=FALSE}
-vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){
+```{r out.width="50%", fig.show="hold"}
+vcf.tbl.g   <- dplyr::filter(variants.gatk,   gt_AR != 0, !is.na(gt_AR))
 
-  vcfR.tid.tmp <- 
-    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), function(x){return(length(x))}))) %>%
-    dplyr::mutate(nb.allele=nb.allele-1)
+# distribution de la couverture des positions variantes
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(x=gt_DP, color=Indiv)) + xlim(0,50) +
+  ylab("DP of variants") + geom_vline(xintercept = DP.min, color = "red") + ggtitle("Variant depth of coverage distribution")
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
 
-  vcfR.tid.list <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){
+print(p)
 
-    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
-      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = ALT,   into = paste0("ALT" ,1:x), sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = AC,    into = paste0("AC" ,1:x),  sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = AF,    into = paste0("AF" ,1:x),  sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = MLEAF, into = paste0("MLEAF" ,1:x),  sep = ",", remove = FALSE) %>%
-      tidyr::separate(col = MLEAC, into = paste0("MLEAC" ,1:x),  sep = ",", remove = FALSE)
+# Distribution de l'allèle ratio
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(gt_AR, color=Indiv)) + 
+  xlab("Alternative allele ratio") + xlim(0,1) + geom_vline(xintercept = AR.min, color = "red")+ ggtitle("Variant allele ratio distribution")
 
-    tmp.list <- lapply(1:x, FUN = function(i){
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
 
-        tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
-                                         gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
-                                         ID_old    = ID,
-                                         gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
-                                         ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
-                                         AC_old    = AC,    AC    = get(paste0("AC", i)),
-                                         AF_old    = AF,    AF    = get(paste0("AF", i)),
-                                         MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
-                                         MLEAC_old = MLEAC, MLEAC = get(paste0("MLEAC", i)),
-                                         ID = paste(CHROM, POS, REF, ALT, sep="_"),
-                                   ) %>%
-        dplyr::select(-gt_ref_AD, 
-                      -paste0("ALT", 1:x), 
-                      -paste0("gt_alt_AD", 1:x),  
-                      -paste0("AC", 1:x),
-                      -paste0("AF", 1:x),
-                      -paste0("MLEAF", 1:x),
-                      -paste0("MLEAC", 1:x)) %>%
-        dplyr::filter(ALT != "*")
+print(p)
 
-        })
-
-        return(tmp.list %>% purrr::reduce(rbind))
-    })
-
-    return(purrr::reduce(vcfR.tid.list, rbind))
-}
 ```
 
-```{r , echo=FALSE, warning=FALSE, error=FALSE}
-haplotype_caller<-paste0(workspace,"/03_snv_indels_calling/",config_ad$project_name,"_snv_indel.vcf")
-vcfR.inst <- vcfR::read.vcfR(haplotype_caller)
+```{r out.width="60%", fig.show="hold"}
+NA_ID <- filter(variants.gatk, nb_NA > 0)$ID
 
-vcfR.tid.list <- vcfR::vcfR2tidy(vcfR.inst)
+variants.gatk %>% filter(ID %in% NA_ID) %>% 
+  ggplot() +  geom_bar(aes(x=as.factor(nb_NA))) + 
+  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("nb of samples") + 
+  ylab("Non covered positions") + ggtitle("Non covered variant positions")
 
-vcfR.tid <- full_join(vcfR.tid.list$fix, vcfR.tid.list$gt, by = c("ChromKey", "POS")) %>% 
-            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))
+```
 
-variants <- vcfR_split_multiAlt(vcfR.tid) %>%
-  dplyr::mutate(gt_AR_corr = dplyr::if_else(gt_AR < AR.min, 0, gt_AR),
-         gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
-         gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)),
-         var_type   = dplyr::if_else(str_length(REF) == str_length(ALT), "SNV", dplyr::if_else(str_length(REF) > str_length(ALT), "DEL", "INS")),
-         var_length = str_length(ALT) - str_length(REF),
-         var_class  = dplyr::if_else(var_length == 0, var_type, 
-                      dplyr::if_else(abs(var_length) == 1,  paste0(var_type, "1"),
-                      dplyr::if_else(abs(var_length) >  50, paste0(var_type, ">50"), paste0(var_type, "<50")))))
+### Low quality variant filtering
 
-variants.control <- variants %>% filter(Indiv %in% dplyr::filter(sample_sheet, condition == "WT")$plant, gt_AR_corr > 0) 
-variants.freq_corr <- variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr)) %>% 
-  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(ID, target,amplicon) %>% count(name="freq")
+* score de qualité : > `r qual.min`
+* nombre de reads qui couvrent la position variant : > `r DP.min`
+* nombre de reads qui portent l'allèle alternatif : > 1
+* min AR : `r AR.min`
+* retirer les positions avec au moins un échantillon non couvert.
 
-variants <- variants %>% dplyr::filter(!ID %in% variants.control$ID) %>% dplyr::full_join(variants.freq_corr, by = c("ID", "target","amplicon"))
-#DT::datatable(variants)
+```{r}
+# remove 
+variants.gatk.flt <- variants.gatk %>% 
+  dplyr::mutate(gt_alt_AD = dplyr::if_else(as.numeric(gt_alt_AD) < 2, 0, as.numeric(gt_alt_AD)),
+                gt_AR     = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP)) %>% 
+  dplyr::filter(!is.na(gt_DP), gt_DP >= DP.min, QUAL >= qual.min, nb_NA == 0, freq_ho == freq)
+
 ```
+
+## Descriptive stat of variants
+
+### variant count per sample
+
+```{r }
+
+vcf.tbl.q.g <- filter(variants.gatk.flt, gt_AR != 0, !is.na(gt_AR))
+
+vcf.tbl.q.g %>% 
+  dplyr::mutate(tag=if_else(freq == length(unique(sample_sheet$sample)), "common", "noCommon")) %>% 
+  dplyr::select(ID, Indiv, var_type, tag) %>%
+  dplyr::mutate(Indiv=if_else(tag == "common", "common", Indiv)) %>% unique() %>%
+  ggplot() + geom_bar(aes(x=Indiv, fill=var_type)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
+  xlab("") + ylab("nb of variants")
+
+```
+* common : les variants communs à tous les échantillons.
+
+### Variant frequency in samples
+
+```{r }
+
+p <- vcf.tbl.q.g %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
+  dplyr::count(name = "nb_var")  %>% filter(freq != length(unique(sample_sheet$sample))) %>% 
+  ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")
+
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
+
+  p <- p + theme(legend.position="none")
+}
+print(p)
+
+# warning -> samples with the same variants or pas de variant
+```
+
+### Substitution
+
+```{r}
+
+p <- vcf.tbl.q.g %>% dplyr::filter(var_type == "SNV", str_length(ALT) == 1) %>% 
+  dplyr::select(ALT, REF, Indiv) %>% group_by(ALT, REF, Indiv) %>% 
+  dplyr::count(name = "nb_sub") %>% 
+  ggplot(aes(x=paste(REF, ALT, sep=">"), y=nb_sub)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + 
+  facet_grid(~paste0("REF : ",REF), scales = "free") + labs(x="", y="nb of variants", color="samples")
+
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
+
+  p <- p + theme(legend.position="none")
+}
+print(p)
+
+```
+
+### Clustering per condition
+
+```{r}
+vcf.tbl.q.g.condition <- vcf.tbl.q.g %>% 
+  dplyr::full_join(dplyr::select(sample_sheet, sample, condition), by = c("Indiv" = "sample")) %>% 
+  tidyr::separate(ANN, c("tmp1", "effect", "tmp2", "gene"), sep = "[|]", extra = "drop") %>% dplyr::select(-tmp1, -tmp2)
+
+tb <- vcf.tbl.q.g.condition %>%  
+  dplyr::select(ID, condition) %>% 
+  dplyr::group_by(ID, condition) %>% unique() %>% 
+  reshape2::dcast(ID~condition, value.var = "condition") %>%
+  dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)})
+
+venn::venn(tb[,-1])
+```
+
+### Clustering per sample
+
+```{r }
+
+df <- vcf.tbl.q.g %>% dplyr::filter( freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv , value.var = "gt_AR")
+
+# calcul la distance entre les lignes de la matrice
+# vcf.tidy.bin[,-1] : sans la 1ere colonne qui contient les IDs
+dist  <- dist(t(df[,-1]), method = "euclidean")
+
+# clustering hiérarchique
+hc    <- hclust(dist)
+
+# arranger la structure du dendrogramme
+dend  <- as.dendrogram(hc)
+
+col.rainbow <- colorspace::rainbow_hcl(length(unique(sample_sheet$condition)))
+
+col <- col.rainbow[as.factor(sample_sheet$condition)]
+names(col) <- sample_sheet$sample
+
+labels_colors(dend) <- col[order.dendrogram(dend)]
+# dessiner le dendrogramme
+plot(dend, main = "Sample clustering using all variants")
+```
+
+```{r, eval=FALSE}
+## annotations
+
+colname.ANN <- c("Allele", "effect", "impact", "GeneName", 
+                 "GeneID", "FeatureType", "FeatureID", "TranscriptBiotype", 
+                 "Rank", "HGVS.c", "HGVS.p", "cDNA_len:", 
+                 "CDS_len", "Protein_len", "DistanceToFeature", "Warnings")
+
+exlude_effect <- c("upstream_gene_variant", "downstream_gene_variant", "intergenic_region", "intragenic_variant", "intragenic_variant", 
+                   "3_prime_UTR_variant", "", "", "")
+
+ID.ANN <- vcf.tbl.q.g %>% dplyr::select(ID, ANN) %>% unique()
+
+ANN.split <- lapply(1:dim(ID.ANN)[1], FUN = function(x){
+
+  tmp <- ID.ANN$ANN[x] %>% str_split(",") %>% unlist()
+  data.frame(ID = rep(ID.ANN$ID[x], length(tmp)), ANN = tmp)
+
+}) %>% purrr::reduce(rbind) %>% 
+  tidyr::separate(col = ANN, into = colname.ANN, sep = "[|]") %>% 
+  dplyr::filter(!effect %in% exlude_effect) %>% 
+  dplyr::select(ID, effect) %>% unique()
+
+effect.split <- lapply(1:dim(ANN.split)[1], function(x){
+
+  tmp <- ANN.split$effect[x] %>% stringr::str_split("&") %>% unlist()
+  data.frame(ID = rep(ANN.split$ID[x], length(tmp)), effect = tmp)
+}) %>% purrr::reduce(rbind)
+
+effect.split %>% dplyr::group_by(effect) %>% dplyr::count()
+
+```

### 2023-07-07 script/report.Rmd
debug report

--- 2023-06-20 script/report.Rmd
@@
     html_document:
+        highlight: tango
         number_sections: yes
+        theme: default
+        toc: yes
+        toc_depth: 3
+        toc_float:
+            collapsed: yes
+            smooth_scroll: yes
 params:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE)
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
 ```
@@
 # title: "`r params$title`"
-# workspace<-params$result_dir
-# sample_sheet.file <- params$sample_sheet
-# DP.min <- params$DP.min
-# AR.min <- params$AR.min
-# qual.min <- params$qual.min
+workspace <-params$result_dir
+sample_sheet.file <- params$sample_sheet
+DP.min <- params$DP.min
+AR.min <- params$AR.min
+qual.min <- params$qual.min
+projectName <- params$projectName
+#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
+#projectName <- "PE-WGS"
 
-projectName <- "PE-WGS"
-
-workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
+#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
 #workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
 
-sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
+#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
 #sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
 
-DP.min <- 4
-AR.min <- 0.8
-qual.min <- 30
-
-# results
-trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
-vcf.file <- paste0(workspace,"/03_snv_indels_calling/", projectName , "_snv_indel.vcf")
-sv.file  <- paste0(workspace,"/04_sv_calling/", projectName , "_sv.vcf")
-
-if(!file.exists(sample_sheet.file)) stop("Error: ")
-if(!file.exists(trim.file)) stop("Error: ")
-if(!file.exists(vcf.file)) stop("Error: ")
-
-eval.SV <- file.exists(sv.file)
-
+# DP.min <- 4
+# AR.min <- 0.8
+# qual.min <- 30
+nbr.alt.min <- 3
 ```
 
-```{r }
+```{r}
+#libraries
 library(DT)
@@
 library(ComplexHeatmap)
+library(dendextend) 
+library(GenomicRanges)
 library(gplots)
 source("functions.R")
+```
+
+# Objective
+
+# Input data
+## Genome version
+```{r, eval=FALSE}
+print(basename(Genome.file))
+```
+
+## Samples
+```{r}
+# error message if sample sheet not exist
+if(!file.exists(sample_sheet.file)) stop("Error: ")
+
+# read sample sheet
+sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
+```
+
+```{r}
+sampleNames <- unique(sample_sheet$sample)
+```
+
+```
+List of samples : `r sampleNames`
+```
+
+```{r list samples}
+
+dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
+  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe', caption = "List of conditions")
 
@@
 
-# Samples
-```{r}
+# MsGenoVa workflow
 
-sample_sheet <- read.table(file = sample_sheet.file, header = TRUE, sep = "\t")
-
-unique(sample_sheet$sample)
-```
-List of samples : `r unique(sample_sheet$sample)`
-
-```{r}
-
-dplyr::select(sample_sheet, sample, condition) %>% unique() %>% dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
-  knitr::kable(align = "lccrr")
- # DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
-
-```
-
-<!-- # Quality control -->
-
-```{r quality_control}
-
-  out <- NULL
-
-  out <- c(out, knitr::knit_child("quality_control.Rmd"), quiet=TRUE)
-
-  knitr::asis_output(out)
-
-```
-
-<!-- # SNV and Small InDels Detection -->
-
-```{r SNV_InDels_calling}
-
-  out <- NULL
-
-  out <- c(out, knitr::knit_child("SNV_InDels_calling.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
-
-```
-
-<!-- # SV Detection -->
-
-```{r SV_calling, eval=eval.SV}
-
-  out <- NULL
-
-  out <- c(out, knitr::knit_child("SV_calling.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
-
-```
-
-```{r tDNA_scan, eval=FALSE, include=FALSE}
-
-  out <- NULL
-
-  out <- c(out, knitr::knit_child("tDNA_scan.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
-
-```
-
-```{r tools, eval=FALSE}
-
-  out <- NULL
-
-  out <- c(out, knitr::knit_child("tools.Rmd"), quiet=TRUE)
-
-knitr::asis_output(out)
-
-```
+![alt text here](workflow.png)

### 2023-07-07 script/report.Rmd
update report

--- 2023-07-06 script/report.Rmd
@@
 
+```{r}
+# title: "`r params$title`"
+workspace <-params$result_dir
+sample_sheet.file <- params$sample_sheet
+DP.min <- params$DP.min
+AR.min <- params$AR.min
+qual.min <- params$qual.min
+projectName <- params$projectName
+#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
+#projectName <- "PE-WGS"
+
+#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
+#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
+
+#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
+#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+
+# DP.min <- 4
+# AR.min <- 0.8
+# qual.min <- 30
+nbr.alt.min <- 3
+```
+
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold", eval = FALSE)
 ```
@@
 library(tidyr)
-library(yaml)
 library(stringr)
-library(ComplexHeatmap)
 library(dendextend) 
-library(GenomicRanges)
-library(gplots)
+#library(GenomicRanges)
 source("functions.R")
@@
 ```{r}
-vcf.file <- paste0(workspace,"/03_snv_indels_calling/", projectName , "_snv_indel.vcf")
+vcf.file <- list.files(path = paste0(workspace,"/03_snv_indels_calling"), pattern = "_snv_indel.vcf" , full.names = T)
 if(!file.exists(vcf.file)) stop("Error: ")

### 2023-07-07 script/SNV_InDels_calling.Rmd

@@
-# SNV and small Indels
+---
+title: "Report"
+author: " "
+date: "2022-08-16"
+output:
+    html_document:
+        highlight: tango
+        number_sections: yes
+        theme: default
+        toc: yes
+        toc_depth: 3
+        toc_float:
+            collapsed: yes
+            smooth_scroll: yes
+params:
+  DP.min: DP.min
+  AR.min: AR.min
+  qual.min: qual.min
+  result_dir: result_dir
+  sample_sheet: sample_sheet
+  rmd: "report.Rmd"
+  AR: AR
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
+```
+
+```{r}
+# title: "`r params$title`"
+# workspace<-params$result_dir
+# sample_sheet.file <- params$sample_sheet
+# DP.min <- params$DP.min
+# AR.min <- params$AR.min
+# qual.min <- params$qual.min
+
+# Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
+Genome.file <- "Physcomitrium_patens_V4_genome.fasta"
+
+projectName <- "PE-WGS_v4"
+
+workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS_v4/results"
+#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
+
+sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS_v4/sample_sheet_PE.txt"
+#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"
+
+DP.min <- 4
+AR.min <- 0.8
+qual.min <- 30
+nbr.alt.min <- 3
+```
+
+```{r}
+#libraries
+library(DT)
+library(ggplot2)
+library(vcfR)
+library(dplyr)
+library(tidyr)
+library(yaml)
+library(stringr)
+library(ComplexHeatmap)
+library(dendextend) 
+library(GenomicRanges)
+library(gplots)
+source("functions.R")
+```
+
+# Input data
+## Genome version
+```{r}
+print(basename(Genome.file))
+```
+
+## Samples
+```{r}
+# error message if sample sheet not exist
+if(!file.exists(sample_sheet.file)) stop("Error: ")
+
+# read sample sheet
+sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
+```
+
+```{r}
+sampleNames <- unique(sample_sheet$sample)
+```
+
+```
+List of samples : `r sampleNames`
+```
+
+```{r list samples}
+
+dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
+  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe', caption = "List of conditions")
+
+```
+
+# MsGenoVa workflow
+
+![alt text here](workflow.png)
+
+# Sequence trimming step
+
+Sequences were trimmed with Trimmomatic tools...
+
+```{r trimming stat}
+trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
+if(!file.exists(trim.file)) stop("Error: ")
+
+read.table(text = gsub(" ","\t",readLines(trim.file)),  header=TRUE) %>%
+  DT::datatable(rownames = FALSE, caption = "Read cleaning results",
+                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe')
+
+```
+
+# Mapping step
+
+Sequences were aligned using BWA MEM (v?) ...
+
+```{r mapping step}
+map.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(!file.exists(map.file)) stop("Error: ")
+```
+
+## Mapping rate
+
+```{r mapping rate}
+flagstat.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
+if(file.exists(flagstat.file)){
+    flagstat <- read.table(flagstat.file, sep="\t", header=FALSE) %>%
+      dplyr::mutate(mapped.reads=V3-V4, mapped.read.ratio=round(mapped.reads/(V2)*100,2)) %>% dplyr::select(-V3, -V4)
+    names(flagstat) <- c("sample", "raw.reads", "mapped.reads", "mapped.read.ratio")
+
+  }
+
+#count
+nbr.reads.melt <- flagstat %>% dplyr::mutate(xlabel=paste0(sample, "(", mapped.read.ratio, "%)"), unmapped.reads=raw.reads-mapped.reads) %>% 
+  dplyr::select(mapped.reads, unmapped.reads, xlabel) %>%
+  reshape2::melt(id = "xlabel", value.name = "nbr.reads")
+
+```
+
+```{r mapping rate plot}
+p <- ggplot(data = nbr.reads.melt) + geom_col(aes(y=xlabel, x=nbr.reads, fill=variable)) + 
+  labs(x="nb of reads", y="samples (mapping rate %)", fill="")
+
+print(p)
+```
+
+```{r mapping rate summary}
+DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+              class = 'cell-border stripe', rownames = FALSE, caption = "")
+```
+
+## Mapping coverage
+
+```{r mapping coverage}
+cov.list <- lapply(sample_sheet$sample, function(x){
+  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
+  if(file.exists(file)){
+
+    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
+  }
+})
+
+cov <- purrr::reduce(cov.list, rbind)
+colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage(%)",	"meandepth(X)",	"meanbaseq",	"meanmapq", "sample")
+cov$coverage <- round(cov$coverage, 2)
+
+cov <- cov %>% mutate(xlabel=paste0(rname, "(", endpos-startpos, ")")) %>% 
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "scaffold")) %>%
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "chloroplast")) %>%
+  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "mitochondria"))
+```
+
+```{r mapping coverage plot}
+p <- select(cov, rname, `meandepth(X)`, `coverage(%)`, sample, xlabel) %>% reshape2::melt(id = c("rname", "sample", "xlabel")) %>% 
+  ggplot() + geom_boxplot(aes(x=value, y=xlabel)) + 
+  xlab("") + ylab("chromosome names (size bp)") + 
+  facet_grid(~variable, scales = "free")
+
+print(p)
+```
+```{r mapping coverage summary}
+cov %>% dplyr::select(sample, rname, `meandepth(X)`, `coverage(%)`) %>% 
+  DT::datatable(rownames = FALSE, caption = "",
+                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
+                class = 'cell-border stripe')
+```
+
+# Small variant calling step (GATK)
+```{r}
+vcf.file <- paste0(workspace,"/03_snv_indels_calling/", projectName , "_snv_indel.vcf")
+if(!file.exists(vcf.file)) stop("Error: ")
+```
 
@@
 
-vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE)
+vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE, info_fields = c("AC","MQ", "ANN"), 
+                                                     format_fields = c("Indiv", "AD", "DP"))
 
-vcfR.tb <- full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS")) %>% 
-  dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))
+vcfR.tb <- dplyr::full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS"))
+rm(vcfR.tb.list)
 
 # split multiallelic variants + add columns  with var type information
-variants.gatk <- vcfR_split_multiAlt(vcfR.tb) %>% 
-  dplyr::mutate(var_length = str_length(ALT) - str_length(REF),
-                var_type   = dplyr::if_else(var_length == 0, "SNV", 
-                             dplyr::if_else(var_length  < 0, "DEL", "INS")),
-                ID = paste(CHROM, POS, var_type, var_length, sep = "_"),
-                tools = "GATK")
+variants.gatk <- vcfR_split_multiAlt(vcfR.tb, nbrAlt = 2)
 
-#!!!!!! CA    AA   
+variants.gatk <- variants.gatk %>%
+  tidyr::separate(gt_AD, c("gt_ref_AD", "gt_alt_AD"), sep = ",") %>%
+  dplyr::mutate(gt_AR = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP),
+                var_length = str_length(ALT) - str_length(REF) + 1,
+                var_type   = dplyr::if_else(var_length == 1, "SNV", 
+                             dplyr::if_else(var_length  < 1, "DEL", "INS")),
+                ID = paste(CHROM, POS, REF, ALT, sep = "_"), tools = "GATK")
+
+variants.gatk <- variant.freq(variants.gatk)
+
 ```
 
-## Quality control and filtering
+## Variant calling quality control
 
-> Before Filtering
+### Score quality
 
-```{r score quality, out.width="50%", fig.show="hold", message=FALSE, warning=FALSE, echo=FALSE}
-
+```{r out.width="60%", fig.show="hold"}
 # Score de qualité
@@
 
+```
+
+### Variant coverage
+
+```{r out.width="50%", fig.show="hold"}
 vcf.tbl.g   <- dplyr::filter(variants.gatk,   gt_AR != 0, !is.na(gt_AR))
@@
 
-# compter le nombre de positions non couvertes par échantillon.
-p <- ggplot(data = vcf.tbl.g) +  geom_bar(aes(x=Indiv)) + 
-  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("") + 
-  ylab("Non covered positions")+ ggtitle("Non covered variant positions")
-print(p)
-
 ```
 
-```{r var depth, out.width="50%", fig.show="hold"}
+```{r out.width="60%", fig.show="hold"}
 NA_ID <- filter(variants.gatk, nb_NA > 0)$ID
@@
   ggplot() +  geom_bar(aes(x=as.factor(nb_NA))) + 
-  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("") + 
+  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("nb of samples") + 
   ylab("Non covered positions") + ggtitle("Non covered variant positions")
@@
 
-> filtering threshold
+### Low quality variant filtering
 
@@
 
-```{r qc filtre}
-
-# filter variants with low quality/coverage + add amplicon positions
-variants.gatk <- filter(variants.gatk, nb_NA == 0)
-variants.gatk.flt <- qc_filtering(variants.gatk, DP.min, AR.min, qual.min)
-
-# dplyr::select(CHROM, POS, ID, REF, ALT, QUAL, EFF, Indiv, gt_GT, gt_AD, gt_DP, gt_AR, 
-#               var_length, var_type,  amplicon, locus, strand, freq, tools)
+```{r}
+# remove 
+variants.gatk.flt <- variants.gatk %>% 
+  dplyr::mutate(gt_alt_AD = dplyr::if_else(as.numeric(gt_alt_AD) < 2, 0, as.numeric(gt_alt_AD)),
+                gt_AR     = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP)) %>% 
+  dplyr::filter(!is.na(gt_DP), gt_DP >= DP.min, QUAL >= qual.min, nb_NA == 0, freq_ho == freq)
 
@@
 
-## Descriptive statistics
+## Descriptive stat of variants
 
-### Number of small variants
+### variant count per sample
 
-```{r nb_SNV, out.width="90%"}
+```{r }
 
 vcf.tbl.q.g <- filter(variants.gatk.flt, gt_AR != 0, !is.na(gt_AR))
-vcf.tbl.q.g %>% dplyr::filter(freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv, var_type) %>% unique() %>% 
+
+vcf.tbl.q.g %>% 
+  dplyr::mutate(tag=if_else(freq == length(unique(sample_sheet$sample)), "common", "noCommon")) %>% 
+  dplyr::select(ID, Indiv, var_type, tag) %>%
+  dplyr::mutate(Indiv=if_else(tag == "common", "common", Indiv)) %>% unique() %>%
   ggplot() + geom_bar(aes(x=Indiv, fill=var_type)) + 
@@
   xlab("") + ylab("nb of variants")
+
 ```
+* common : les variants communs à tous les échantillons.
 
-### Variant frequency
+### Variant frequency in samples
 
-```{r SNV_freq, out.width="90%"}
+```{r }
 
 p <- vcf.tbl.q.g %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
-  dplyr::filter(freq != length(unique(sample_sheet$sample))) %>% dplyr::count(name = "nb_var") %>% 
+  dplyr::count(name = "nb_var")  %>% filter(freq != length(unique(sample_sheet$sample))) %>% 
   ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
@@
 
-vcf.tbl.g %>% filter(POS == 1548127)
-
 # warning -> samples with the same variants or pas de variant
@@
 
-### Clustering
+## Specific variants
 
-```{r SNV_clustering, out.width="90%"}
+```{r}
+vcf.tbl.q.g.condition <- vcf.tbl.q.g %>% 
+  dplyr::full_join(dplyr::select(sample_sheet, sample, condition), by = c("Indiv" = "sample")) %>% 
+  tidyr::separate(ANN, c("tmp1", "effect", "tmp2", "gene"), sep = "[|]", extra = "drop") %>% dplyr::select(-tmp1, -tmp2)
 
-df <- vcf.tbl.q.g %>% dplyr::filter( freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
-  reshape2::dcast(ID~Indiv , value.var = "gt_AR")
+tb <- vcf.tbl.q.g.condition %>%  
+  dplyr::select(ID, condition) %>% 
+  dplyr::group_by(ID, condition) %>% unique() %>% 
+  reshape2::dcast(ID~condition, value.var = "condition") %>%
+  dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)})
 
-# calcul la distance entre les lignes de la matrice
-# vcf.tidy.bin[,-1] : sans la 1ere colonne qui contient les IDs
-dist  <- dist(t(df[,-1]), method = "euclidean")
-
-# clustering hiérarchique
-hc    <- hclust(dist)
-
-# arranger la structure du dendrogramme
-dend  <- as.dendrogram(hc)
-
-# dessiner le dendrogramme
-plot(dend, main = "Sample clustering using all variants")
+venn::venn(tb[,-1])
 ```
 
-### 3 type de substitution for AG > TA ...
+On va s'interesser aux :
+
+* variants specifiques à chaques groupes.
+* variants communs uniquement entre les groups **PE** et **SPE**.
 
@@
 
-p <- vcf.tbl.q.g %>% dplyr::filter(var_type == "SNV", str_length(ALT) == 1) %>% 
-  dplyr::select(ALT, REF, Indiv) %>% group_by(ALT, REF, Indiv) %>% 
-  dplyr::count(name = "nb_sub") %>% 
-  ggplot(aes(x=paste(REF, ALT, sep=">"), y=nb_sub)) + geom_boxplot(outlier.shape = NA) +
-  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2))
+# je retirer les variations communs
+
+freq_condition <- lapply(unique(vcf.tbl.q.g.condition$condition), function(x){
+
+  vcf.tbl.q.g.condition %>% filter(condition == x) %>% group_by(ID) %>% dplyr::count(name = paste0("freq_", x)) %>% return()
+}) %>% purrr::reduce(full_join, by = "ID") %>% dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, x)})
+
+vcf.tbl.q.g <- vcf.tbl.q.g.condition %>% full_join(freq_condition, by = "ID")
+
+```
+
+### in WT group
+
+```{r, out.width="70%", fig.show="hold"}
+
+gatk_var_WT  <- filter(vcf.tbl.q.g, freq_WT  == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+gatk_var_WT %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+gatk_var_WT %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_ref_AD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+
+names(gatk_var_WT)
+```
+
+### in Pro group
+
+```{r, out.width="70%", fig.show="hold"}
+# 
+gatk_var_Pro <- filter(vcf.tbl.q.g, freq_Pro == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+gatk_var_Pro %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+gatk_var_Pro %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_ref_AD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in T group
+
+```{r, out.width="70%", fig.show="hold"}
+gatk_var_T   <- filter(vcf.tbl.q.g, freq_T   == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+gatk_var_T %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+gatk_var_T %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_ref_AD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in PE group
+
+```{r, out.width="70%", fig.show="hold"}
+gatk_var_PE  <- filter(vcf.tbl.q.g, freq_PE  == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+gatk_var_PE %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+gatk_var_PE %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_ref_AD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in SPE group
+
+```{r, out.width="70%", fig.show="hold"}
+gatk_var_SPE <- filter(vcf.tbl.q.g, freq_SPE == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+gatk_var_SPE %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+gatk_var_SPE %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_ref_AD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in PE, SPE groups
+
+```{r, out.width="70%", fig.show="hold"}
+
+gatk_var_PE_SPE <- filter(vcf.tbl.q.g, freq_SPE + freq_PE == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+gatk_var_PE_SPE %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+gatk_var_PE_SPE %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_ref_AD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+# Small variant calling step (Varscan)
+
+```{r}
+
+snv_varscan.vcfR.tb <- lapply(sample_sheet$sample, function(x){
+  file <- paste0(workspace,"/03_bis_small_indel_varscan/", x, ".vcf.gz")
+  if(file.exists(file)){
+
+    vcfR <- vcfR::read.vcfR(file, verbose = FALSE)
+    vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE, info_fields = c("AC","MQ", "ANN"), 
+                                                         format_fields = c("Indiv", "DP", "RD", "AD"))
+
+    vcfR.tb <- dplyr::full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS"))
+
+    return(vcfR.tb)
+  }
+}) %>% purrr::reduce(rbind)
+
+# filter(snv_varscan.vcfR.tb, CHROM == "Chr08", Indiv == "PE-1") #, str_detect(POS,"10812105"))
+
+```
+
+```{r}
+
+variants.varscan <- snv_varscan.vcfR.tb %>%
+  dplyr::mutate(gt_AR = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, gt_AD/gt_DP),
+                var_length = str_length(ALT) - str_length(REF) + 1,
+                var_type   = dplyr::if_else(var_length == 1, "SNV", 
+                             dplyr::if_else(var_length  < 1, "DEL", "INS")),
+                ID = paste(CHROM, POS, REF, ALT, sep = "_"), tools = "Varscan")
+
+variants.varscan <- variant.freq(variants.varscan)
+
+```
+
+## Variant calling quality control
+
+### Variant coverage
+
+```{r out.width="50%", fig.show="hold"}
+vcf.tbl.g   <- dplyr::filter(variants.varscan,   gt_AR != 0, !is.na(gt_AR))
+
+# distribution de la couverture des positions variantes
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(x=gt_DP, color=Indiv)) + xlim(0,50) +
+  ylab("DP of variants") + geom_vline(xintercept = DP.min, color = "red") + ggtitle("Variant depth of coverage distribution")
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
+
+print(p)
+
+# Distribution de l'allèle ratio
+p <- ggplot(data=vcf.tbl.g) + geom_density(aes(gt_AR, color=Indiv)) + 
+  xlab("Alternative allele ratio") + xlim(0,1) + geom_vline(xintercept = AR.min, color = "red")+ ggtitle("Variant allele ratio distribution")
+
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
+
+print(p)
+```
+
+### Low quality variant filtering
+
+* nombre de reads qui couvrent la position variant : > `r DP.min`
+* nombre de reads qui portent l'allèle alternatif : > 1
+* min AR : `r AR.min`
+
+```{r}
+# remove 
+variants.varscan.flt <- variants.varscan %>% 
+  dplyr::mutate(gt_AD = dplyr::if_else(gt_AD < 2, 0, as.numeric(gt_AD)),
+                gt_AR = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, gt_AD/gt_DP)) %>% 
+  dplyr::filter(!is.na(gt_DP), gt_DP >= DP.min, FILTER == "PASS", freq_ho == freq)
+
+```
+
+## Descriptive stat of variants
+
+### variant count per sample
+
+```{r }
+
+vcf.tbl.q.g <- filter(variants.varscan.flt, gt_AR != 0, !is.na(gt_AR))
+
+vcf.tbl.q.g %>% 
+  dplyr::mutate(tag=if_else(freq == length(unique(sample_sheet$sample)), "common", "noCommon")) %>% 
+  dplyr::select(ID, Indiv, var_type, tag) %>%
+  dplyr::mutate(Indiv=if_else(tag == "common", "common", Indiv)) %>% unique() %>%
+  ggplot() + geom_bar(aes(x=Indiv, fill=var_type)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
+  xlab("") + ylab("nb of variants")
+
+```
+
+### Variant frequency in samples
+
+```{r }
+
+p <- vcf.tbl.q.g %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
+  dplyr::count(name = "nb_var")  %>% filter(freq != length(unique(sample_sheet$sample))) %>% 
+  ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")
 
@@
 
+# warning -> samples with the same variants or pas de variant
 ```
 
-### 4 annotations
+## Specific variants
+
 ```{r}
+vcf.tbl.q.g.condition <- dplyr::full_join(vcf.tbl.q.g, dplyr::select(sample_sheet, sample, condition), by = c("Indiv" = "sample"))
+
+tb <- vcf.tbl.q.g.condition %>%  
+  dplyr::select(ID, condition) %>% 
+  dplyr::group_by(ID, condition) %>% unique() %>% 
+  reshape2::dcast(ID~condition, value.var = "condition") %>%
+  dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)})
+
+venn::venn(tb[,-1])
+```
+
+On va s'interesser aux :
+
+* variants specifiques à chaques groupes.
+* variants communs uniquement entre les groups **PE** et **SPE**.
+
+```{r}
+
+freq_condition <- lapply(unique(vcf.tbl.q.g.condition$condition), function(x){
+
+  vcf.tbl.q.g.condition %>% filter(condition == x) %>% group_by(ID) %>% dplyr::count(name = paste0("freq_", x)) %>% return()
+}) %>% purrr::reduce(full_join, by = "ID") %>% dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, x)})
+
+vcf.tbl.q.g <- vcf.tbl.q.g.condition %>% full_join(freq_condition, by = "ID")
+
+```
+
+### in WT group
+
+```{r, out.width="70%", fig.show="hold"}
+# 
+varscan_var_WT  <- filter(vcf.tbl.q.g, freq_WT  == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+varscan_var_WT %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+varscan_var_WT %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_RD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in Pro group
+
+```{r, out.width="70%", fig.show="hold"}
+# 
+varscan_var_Pro <- filter(vcf.tbl.q.g, freq_Pro == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+varscan_var_Pro %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+varscan_var_Pro %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_RD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in T group
+
+```{r, out.width="70%", fig.show="hold"}
+varscan_var_T   <- filter(vcf.tbl.q.g, freq_T   == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+varscan_var_T %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+varscan_var_T %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_RD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in PE group
+
+```{r, out.width="70%", fig.show="hold"}
+varscan_var_PE  <- filter(vcf.tbl.q.g, freq_PE  == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+varscan_var_PE %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+varscan_var_PE %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_RD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in SPE group
+
+```{r, out.width="70%", fig.show="hold"}
+varscan_var_SPE <- filter(vcf.tbl.q.g, freq_SPE == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+varscan_var_SPE %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+varscan_var_SPE %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_RD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+### in PE, SPE groups
+
+```{r, out.width="70%", fig.show="hold"}
+
+varscan_var_PE_SPE <- filter(vcf.tbl.q.g, freq_SPE + freq_PE == freq) %>% 
+  dplyr::select(-freq_WT, -freq_Pro, -freq_T, -freq_PE, -freq_SPE, -ChromKey, -freq_ho)
+
+varscan_var_PE_SPE %>% select(ID, gt_AR, Indiv) %>% 
+  reshape2::dcast(ID~Indiv, value.var = "gt_AR") %>%
+  mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)}) %>% 
+  dplyr::select(-ID) %>% venn::venn()
+
+varscan_var_PE_SPE %>% dplyr::select(-ID, -FILTER, -AC, -MQ, -gt_RD, -gt_GT_alleles, -nb_0, -nb_NA) %>%
+  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'), 
+                class = 'cell-border stripe', caption = "")
+```
+
+## annotations
+```{r, eval=FALSE}
 
@@
 ```
+
+# Structural variant calling step (delly)
+
+```{r, eval=FALSE}
+sv.vcfR.tb <- lapply(sample_sheet$sample, function(x){
+  file <- paste0(workspace,"/04_delly/", x, ".vcf.gz")
+  if(file.exists(file)){
+
+    vcfR <- vcfR::read.vcfR(file, verbose = FALSE)
+    vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE)
+
+    vcfR.tb <- dplyr::full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS")) %>% 
+               dplyr::mutate(SVLEN = dplyr::if_else(SVTYPE %in% c("DUP", "DEL", "INV"), END-POS+1, SVLEN),
+                             ID    = paste(CHROM, POS, REF, SVTYPE, sep = "_"))
+
+    return(vcfR.tb)
+  }
+}) %>% purrr::reduce(rbind)
+
+```
+
+* Quality score
+
+```{r, eval=FALSE}
+
+# distribution de la couverture des positions variantes
+p <- ggplot(data=sv.vcfR.tb) + geom_density(aes(x=QUAL, color=Indiv)) + xlim(0,50) +
+  ylab("DP of variants") + geom_vline(xintercept = DP.min, color = "red") + ggtitle("Variant depth of coverage distribution")
+if (length(unique(vcf.tbl.g$Indiv)) > 15){
+  p <- p + theme(legend.position="none")
+}
+
+print(p)
+```
+
+## Filter low quality variants
+
+```{r, eval=FALSE}
+
+tmp <- dplyr::filter(sv.vcfR.tb, FILTER == "PASS") %>% dplyr::select(CHROM, POS, END, Indiv, ID, SVTYPE)
+
+gr.reduce <- lapply(c("DEL", "DUP", "INV"), function(x){
+
+  tmp.type <- tmp %>% dplyr::filter(SVTYPE == "DEL")
+  gr <- GenomicRanges::GRanges(seqnames=tmp.type$CHROM,
+           ranges=IRanges(start=tmp.type$POS,
+                          end=tmp.type$END))
+
+  reduce(gr, with.revmap=TRUE)
+  })
+names(gr.reduce) <- c("DEL", "DUP", "INV")
+
+gr.bis <- gr[1]
+
+gr[which(gr$SVTYPE == "DUP"),]
+
+subsetByOverlaps(gr, gr.bis)
+
+gr$SVTYPE
+
+gr[,gr$start == 145049]
+
+filter(sv.vcfR.tb, SVTYPE == "INS") %>% arrange(POS) # 29823461
+
+```
+
+```{r, eval=FALSE}
+
+distance <- 0
+
+tmp <- dplyr::filter(sv.vcfR.tb, FILTER == "PASS")
+
+gr.start <- lapply(unique(tmp$SVTYPE), function(x){
+
+  tmp.type <- tmp %>% dplyr::filter(SVTYPE == x)
+  gr <- GenomicRanges::GRanges(seqnames=tmp.type$CHROM,
+           ranges=IRanges(start=tmp.type$POS,
+                          width = 1))
+
+  index <- GenomicRanges::distanceToNearest(gr, select="all") %>% as_tibble() %>% 
+    dplyr::filter(distance == 0) %>% 
+    group_by(queryHits) %>% dplyr::summarise(index = paste0(subjectHits, collapse = ", ")) %>% 
+    dplyr::mutate(vec = paste0("c(",queryHits, ", ", index, ")")) %>% 
+    group_by(vec) %>% summarise(index = paste0(sort(eval(expr = parse(text = vec))), collapse = ",")) %>% 
+    dplyr::select(index) %>% unique()
+
+  lapply(paste0("c(", index$index, ")"), function(x){ tmp.type[eval(parse(text=x)),] %>% mutate(POS.bis = min(POS)) }) %>% purrr::reduce(rbind)
+  })
+names(gr.start) <- unique(tmp$SVTYPE)
+
+gr.bis <- gr[1]
+
+gr[which(gr$SVTYPE == "DUP"),]
+
+subsetByOverlaps(gr, gr.bis)
+
+gr$SVTYPE
+
+gr[,gr$start == 145049]
+
+filter(sv.vcfR.tb, SVTYPE == "INS") %>% arrange(POS) # 29823461
+
+tmp.type[c(5928, 2769, 5929, 2770, 5155),]
+
+```
+
+```{r, eval=FALSE}
+query <- GRanges("A", IRanges(c(2, 5), c(8, 15)))
+query
+subject <- GRanges("A", IRanges(c(1, 4, 10, 1), width = 1))
+subject
+```
+
+```{r, eval=FALSE}
+#distance(query, subject)
+
+nearestKNeighbors(query, subject)
+nearest(query, subject, select="all")
+ distanceToNearest(subject, select="all")
+
+nearestKNeighbors(query, subject, k=3)
+
+nearestKNeighbors(query, subject, select="all")
+```
+
+```{r, eval=FALSE}
+
+sv.vcfR.tb.flt <-
+
+  dplyr::filter(sv.vcfR.tb, FILTER == "PASS") %>% dplyr::select(CHROM, POS, END, Indiv)
+
+```
+
+## Variant count
+
+```{r, eval=FALSE}
+
+sv.vcfR.tb <- sv.vcfR.tb %>% dplyr::filter(FILTER == "PASS")
+
+sv.vcfR.tb %>% dplyr::select(ID, Indiv, SVTYPE) %>% unique() %>% 
+  ggplot() + geom_bar(aes(x=Indiv, fill=SVTYPE)) + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
+  xlab("") + ylab("nb of variants")
+```
+
+```{r , eval=FALSE}
+
+sv.freq <- sv.vcfR.tb %>% dplyr::group_by(ID) %>% dplyr::count(name = "freq")
+
+sv.vcfR.tb <- dplyr::full_join(sv.vcfR.tb, sv.freq, by="ID")
+
+p <- sv.vcfR.tb %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
+  dplyr::count(name = "nb_var")  %>% filter(freq < 14) %>% 
+  ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
+  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")
+
+if (length(unique(vcf.tbl.q.g$Indiv)) > 15){
+
+  p <- p + theme(legend.position="none")
+}
+print(p)
+
+# warning -> samples with the same variants or pas de variant
+```
